---
title: "MACA_plate_dissociation.Rmd"
output: html_document
---

Source the `loader.Rmd` file first.

Load the requisite packages and some additional helper functions.

```{r}
library(Seurat)
library(dplyr)
library(Matrix)
```

Read in genes affected by dissociation

```{r}
dissociation_genes = read.csv('~/genes_affected_by_dissociation_unix.csv', stringsAsFactors = FALSE)$Gene
```


Run PCA on dissociation genes only and plot median of each plate

```{r, echo=FALSE, fig.height=4, fig.width=8}
tiss = RunPCA(tiss, pc.genes=dissociation_genes, do.print=FALSE)
tiss <- ProjectPCA(object = tiss, do.print = FALSE)
PCHeatmap(object = tiss, pc.use = 1:6, cells.use = 500, do.balanced = TRUE, label.columns = FALSE, num.genes = 8)

FetchData(tiss, vars.all = c('plate.barcode.x','tissue', 'PC1', 'PC2', 'color')) %>% 
  group_by(plate.barcode.x, tissue, color) %>%
  summarize(median_PC1 = median(PC1),
            median_PC2 = median(PC2)) %>%
  ggplot(aes(x = median_PC1, y = median_PC2, color = tissue)) + geom_point(aes(color = color)) +
       scale_color_identity(breaks = tissue_colors$color, 
                        labels = tissue_colors$anno_tissue, 
                        guide = "legend") + 
  guides(colour = guide_legend(override.aes = list(size=2)))

ggsave('dissociation_genes_pca_medians.pdf')
```
```{r, fig.height=8, fig.width=8}
dissociation_genes_in_data = dissociation_genes[dissociation_genes %in% rownames(tiss@scale.data)]
length(dissociation_genes_in_data)
tiss@meta.data$total.dissociation = Matrix::colMeans(tiss@data[dissociation_genes_in_data,])

FetchData(tiss, vars.all = c('plate.barcode.x', 'anno_tissue', 'color', 'total.dissociation'))%>%
            ggplot(aes(total.dissociation, ..density.., colour = color)) +
  geom_freqpoly(binwidth = .05) +
  facet_wrap(~ anno_tissue, ncol = 4) +
  scale_color_identity(breaks = tissue_colors$color,
                       labels = tissue_colors$anno_tissue,
                       guide = "legend") +
  ggtitle("Total dissociation")
ggsave('dissociation_genes_tissue_distributions.pdf')
```



```{r}
FetchData(tiss, vars.all = c('plate.barcode.x','tissue', 'PC3', 'PC4', 'color')) %>% 
  group_by(plate.barcode.x, tissue, color) %>%
  summarize(med1 = median(PC3),
            med2 = median(PC4)) %>%
  ggplot(aes(x = med1, y = med2, color = tissue)) + geom_point(aes(color = color)) +
       scale_color_identity(breaks = tissue_colors$color, 
                        labels = tissue_colors$anno_tissue, 
                        guide = "legend") + 
  guides(colour = guide_legend(override.aes = list(size=2)))
```

```{r}
PCElbowPlot(tiss)
```

Example code from Josh
    
    FetchData(tiss, vars.all = c('tSNE_1','tSNE_2', 'color')) %>% 
      ggplot(aes(x = tSNE_1, y = tSNE_2)) + geom_point(aes(color = color), size=0.1) +
       scale_color_identity(breaks = tissue_colors$color, 
                            labels = tissue_colors$anno_tissue, 
                            guide = "legend") + 
      guides(colour = guide_legend(override.aes = list(size=2)))




```{r, echo=FALSE, fig.height=4, fig.width=8}
dissociation_genes = read.csv('~/genes_affected_by_dissociation_unix.csv', stringsAsFactors = FALSE)$Gene
tiss = RunICA(tiss, ic.genes=dissociation_genes, ics.compute = 10, rev.ica = FALSE, print.results=FALSE)
tiss <- ProjectDim(object = tiss, do.print = FALSE, reduction.type = 'ica')
```

```{r, echo=FALSE, fig.height=8, fig.width=8}

ICHeatmap(object = tiss, ic.use = 1:10, cells.use = 500, do.balanced = TRUE, 
          label.columns = FALSE, num.genes = 12)

FetchData(tiss, vars.all = c('plate.barcode.x','tissue', 'IC2', 'IC3', 'color')) %>% 
  group_by(plate.barcode.x, tissue, color) %>%
    summarize(med1 = median(IC2),
            med2 = median(IC3)) %>%
  ggplot(aes(x = med1, y = med2, color = tissue)) + geom_point(aes(color = color)) +
       scale_color_identity(breaks = tissue_colors$color, 
                        labels = tissue_colors$anno_tissue, 
                        guide = "legend") + 
  guides(colour = guide_legend(override.aes = list(size=2)))
```
```{r}
GenePlot(tiss, "Jun", "Hspa1a")
```



```{r}
FetchData(tiss, vars.all = c('plate.barcode.x','tissue', 'IC1', 'IC2', 'color')) %>% 
  group_by(plate.barcode.x, tissue, color) %>%
  summarize(med1 = median(IC1),
            med2 = median(IC2)) %>%
  ggplot(aes(x = med1, y = med2, color = tissue)) + geom_point(aes(color = color)) +
       scale_color_identity(breaks = tissue_colors$color, 
                        labels = tissue_colors$anno_tissue, 
                        guide = "legend") + 
  guides(colour = guide_legend(override.aes = list(size=2)))
```
Run PCA with all genes except dissociation genes
```{r}
length(tiss@var.genes)
is_dissociation_gene = tiss@var.genes %in% dissociation_genes
not_dissociation_gene = ! is_dissociation_gene
non_dissociation_genes = tiss@var.genes[not_dissociation_gene]
length(non_dissociation_genes)
```




```{r}
tiss = RunPCA(tiss, pc.genes=non_dissociation_genes, do.print=FALSE)
tiss <- ProjectPCA(object = tiss, do.print = FALSE)
PCHeatmap(object = tiss, pc.use = 1:3, cells.use = 500, do.balanced = TRUE, label.columns = FALSE, num.genes = 8)

FetchData(tiss, vars.all = c('plate.barcode.x','tissue', 'PC1', 'PC2', 'color')) %>% 
  group_by(plate.barcode.x, tissue, color) %>%
  summarize(med1 = median(PC1),
            med2 = median(PC2)) %>%
  ggplot(aes(x = med1, y = med2, color = tissue)) + geom_point(aes(color = color)) +
       scale_color_identity(breaks = tissue_colors$color, 
                        labels = tissue_colors$anno_tissue, 
                        guide = "legend") + 
  guides(colour = guide_legend(override.aes = list(size=2)))
```
Compare to running on all genes

```{r}
tiss = RunPCA(tiss, do.print=FALSE)
tiss <- ProjectPCA(object = tiss, do.print = FALSE)
PCHeatmap(object = tiss, pc.use = 1:3, cells.use = 500, do.balanced = TRUE, label.columns = FALSE, num.genes = 8)
FetchData(tiss, vars.all = c('plate.barcode.x','tissue', 'PC1', 'PC2', 'color')) %>% 
  group_by(plate.barcode.x, tissue, color) %>%
  summarize(med1 = median(PC1),
            med2 = median(PC2)) %>%
  ggplot(aes(x = med1, y = med2, color = tissue)) + geom_point(aes(color = color)) +
       scale_color_identity(breaks = tissue_colors$color, 
                        labels = tissue_colors$anno_tissue, 
                        guide = "legend") + 
  guides(colour = guide_legend(override.aes = list(size=2)))
```
