---
title: "Diaphragm FACS Notebook"
output: html_notebook
---

Enter the directory of the maca folder on your drive and the name of the tissue you want to analyze.

```{r}
tissue_of_interest = "Diaphragm"
```

Load the requisite packages and some additional helper functions.

```{r}
library(here)
library(useful)
library(Seurat)
library(dplyr)
library(Matrix)
library(ontologyIndex)
cell_ontology = get_ontology('https://raw.githubusercontent.com/obophenotype/cell-ontology/master/cl-basic.obo', extract_tags='everything')

validate_cell_ontology = function(cell_ontology_class){
  in_cell_ontology = sapply(cell_ontology_class, function(x) is.element(x, cell_ontology$name))
  if (!all(in_cell_ontology)) {
    message = paste0('"', cell_ontology_class[!in_cell_ontology], '" is not in the cell ontology
')
    stop(message)
  }
}
convert_to_cell_ontology_id = function(cell_ontology_class){
  return(sapply(cell_ontology_class, function(x) as.vector(cell_ontology$id[cell_ontology$name == x])[1]))
}

save_dir = here('00_data_ingest', 'tissue_robj')
```

Load the plate metadata. Check which plates have been downloaded.

```{r}
plate_metadata_filename = here('00_data_ingest', '00_facs_raw_data', 'metadata_FACS.csv')

plate_metadata <- read.csv(plate_metadata_filename, sep=",", header = TRUE)
colnames(plate_metadata)[1] <- "plate.barcode"
plate_metadata
```

"Diaphragm" is a subtissue of "Muscle" and we will have to filter for this.

```{r}
tissue_plates = filter(plate_metadata, subtissue == tissue_of_interest)[,c('plate.barcode','tissue','subtissue','mouse.sex')]
tissue_plates
```
Load the read count data. 
```{r}
#Load the gene names and set the metadata columns by opening the first file
filename = here('00_data_ingest', '00_facs_raw_data', 'FACS', 'Muscle-counts.csv')

raw.data = read.csv(filename, sep=",", row.names=1)
# raw.data = data.frame(row.names = rownames(raw.data))
corner(raw.data)
```

Get the plate barcode of each individual cell

```{r}
plate.barcodes = lapply(colnames(raw.data), function(x) strsplit(strsplit(x, "_")[[1]][1], '.', fixed=TRUE)[[1]][2])
meta.data = plate_metadata[plate_metadata$plate.barcode %in% plate.barcodes, ]
dim(plate_metadata)
dim(meta.data)
corner(meta.data)
```
Use only cells from "Diaphragm"" plate barcodes.

```{r}
subtissue.plates = filter(tissue_plates, subtissue == tissue_of_interest)
subtissue.plates = plate.barcodes %in% subtissue.plates$plate.barcode
sum(subtissue.plates)

raw.data = raw.data[subtissue.plates]
```

Create per-cell metadata from plate barcode metadata

```{r}
barcode.df = t.data.frame(as.data.frame(plate.barcodes[subtissue.plates]))
head(barcode.df)

rownames(barcode.df) = colnames(raw.data)
colnames(barcode.df) = c('plate.barcode')
head(barcode.df)

rnames = row.names(barcode.df)
meta.data <- merge(barcode.df, plate_metadata, by='plate.barcode', sort = F)
row.names(meta.data) <- rnames

# Sort cells by plate barcode because that's how the data was originally
meta.data = meta.data[order(meta.data$plate.barcode), ]
raw.data = raw.data[, rownames(meta.data)]
head(raw.data)
```


```{r}

erccs <- grep(pattern = "^ERCC-", x = rownames(x = raw.data), value = TRUE)
percent.ercc <- Matrix::colSums(raw.data[erccs, ])/Matrix::colSums(raw.data)
ercc.index <- grep(pattern = "^ERCC-", x = rownames(x = raw.data), value = FALSE)
raw.data <- raw.data[-ercc.index,]

tiss <- CreateSeuratObject(raw.data = raw.data, project = tissue_of_interest, 
                    min.cells = 5, min.genes = 5)

tiss <- AddMetaData(object = tiss, meta.data)
tiss <- AddMetaData(object = tiss, percent.ercc, col.name = "percent.ercc")
# Change default name for sums of counts from nUMI to nReads
colnames(tiss@meta.data)[colnames(tiss@meta.data) == 'nUMI'] <- 'nReads'
```


Calculate percent ribosomal genes.

```{r}
ribo.genes <- grep(pattern = "^Rp[sl][[:digit:]]", x = rownames(x = tiss@data), value = TRUE)

percent.ribo <- Matrix::colSums(tiss@raw.data[ribo.genes, ])/Matrix::colSums(tiss@raw.data)

tiss <- AddMetaData(object = tiss, metadata = percent.ribo, col.name = "percent.ribo")
```

A sanity check: reads vs genes.

```{r}
GenePlot(object = tiss, gene1 = "nReads", gene2 = "nGene")
GenePlot(object = tiss, gene1 = "Pax7", gene2 = "Pax3")
GenePlot(object = tiss, gene1 = "Pax7", gene2 = "Myod1")
```

Filter out cells with few reads and few genes.

```{r}
tiss <- FilterCells(object = tiss, subset.names = c("nGene", "nReads"), 
    low.thresholds = c(500, 50000), high.thresholds = c(25000, 2000000))
```


Normalize the data, then regress out correlation with total reads
```{r}
tiss <- NormalizeData(object = tiss)
tiss <- ScaleData(object = tiss, vars.to.regress = c("nReads", "percent.ribo","Rn45s"))
tiss.back <- tiss
```

```{r}
#tiss <- tiss.back
tiss <- FindVariableGenes(object = tiss, do.plot = TRUE, x.high.cutoff = Inf, y.cutoff = 0.5)
```


Run Principal Component Analysis.
```{r}
tiss <- RunPCA(object = tiss, pc.genes = tiss@var.genes, do.print = TRUE, pcs.print = 1:5, 
    genes.print = 5)
tiss <- ProjectPCA(object = tiss, do.print = FALSE)
```


```{r, echo=FALSE, fig.height=6, fig.width=8}
PCHeatmap(object = tiss, pc.use = 1:9, cells.use = 500, do.balanced = TRUE, label.columns = FALSE, num.genes = 16)
tiss.back <- tiss
```

```{r}
PCElbowPlot(object = tiss)
```


```{r}
n.pcs = 10
#tiss <- tiss.back
tiss <- FindClusters(object = tiss, reduction.type = "pca", dims.use = 1:n.pcs, 
    resolution = 1, print.output = 0, save.SNN = TRUE, force.recalc = TRUE)
```


```{r}
tiss <- RunTSNE(object = tiss, dims.use = 1:n.pcs, seed.use = 10, check_duplicates = F, perplexity=30)
```

```{r}
pt.size = 0.5
# note that you can set do.label=T to help label individual clusters
TSNEPlot(object = tiss, do.label = T, pt.size = pt.size)
```

Check expression of genes of interest.
```{r, fig.height=6, fig.width=8}
genes_to_check = c('Pecam1', 'Ptprc', 'Vcam1', 'Pdgfra')

#FeaturePlot(tiss, genes_to_check, pt.size = 0.5, cols.use=c("yellow", "black"), no.legend = FALSE)
FeaturePlot(tiss, genes_to_check, pt.size = pt.size, no.legend = FALSE)
```


```{r, fig.height=6, fig.width=8}
# SC - subtypes
genes_to_check = c('Vcam1', 'Pax7', 'Pax3', 'Myod1', 'Myf5', 'Cd34', 'Itga7', 'Cd44', 'Calcr' #, 'Myog'
                   )
FeaturePlot(tiss, genes_to_check, pt.size = pt.size)
```

```{r, fig.height=6, fig.width=8}
# CD45 - macrophage subtypes
genes_to_check = c('Ptprc', 
                   'Klrb1b', 'Foxp3', # NK-cells
                   'Cd3d', #T-cell
                   'Cd19', 'Cd79a', 'Cd79b', #differentiated B cells
                   'Vpreb3', #naive B-cells
                   'Itgam', 'Fcer1g', 'C1qa' #macrophages and mast cells
                   #, 'Ly6g6d', 
)
FeaturePlot(tiss, genes_to_check, pt.size = pt.size)
```


```{r}
# CD31 - endothelial subtypes
genes_to_check = c('Pecam1' #, 'Vegfa', 'Cd34', 'Ptprc', 'Icam1' #, 'Dysf' #'Rapgef3', 'Vim', 'Cspg4'
)
FeaturePlot(tiss, genes_to_check, pt.size = pt.size)
```

```{r}
# FAP -  subtypes
genes_to_check = c('Atxn1', 'Pdgfra' #'Dysf', 'Vim', 'Cspg4'
)
FeaturePlot(tiss, genes_to_check, pt.size = pt.size)
```
How big are the clusters?
```{r}
table(tiss@ident)
```

Color by metadata, like plate barcode, to check for batch effects.
```{r}
#TSNEPlot(object = tiss, do.return = TRUE, group.by = "plate.barcode")
TSNEPlot(object = tiss, do.return = TRUE, group.by = "mouse.id", pt.size = pt.size)
#table(as.character(tiss@ident), as.character(tiss@meta.data$plate.barcode))
table(as.character(tiss@ident), as.character(tiss@meta.data$mouse.id))
```

Find differentially expressed markers.
```{r}
tiss.markers <- FindAllMarkers(object = tiss, only.pos = TRUE, min.pct = 0.25, 
    thresh.use = 0.25)
```

Display top 24 markers per cluster.
```{r}
tiss.markers %>% group_by(cluster) %>% top_n(24, avg_diff)
```


```{r}
# To change the y-axis to show raw counts, add use.raw = T.
genes_to_check = c('Pecam1', 'Ptprc', 'Vcam1', 'Pdgfra', 'Cd19')
VlnPlot(tiss, genes_to_check #, pt.size = pt.size
        )
```
Assigning cell type identity to clusters
At a coarse level, we can use canonical markers to match the unbiased clustering to known cell types:
0: alpha 1: beta 2: beta 3: exocrine 4: duct 5: delta 6: gamma 7: endothelial 8: immune 9: stellate
```{r}
# stash current cluster IDs
tiss <- StashIdent(object = tiss, save.name = "cluster.ids")

# enumerate current cluster IDs and the labels for them
cluster.ids <- c(0, 1, 2, 3, 4, 5)
cell_ontology_class <- c("mesenchymal stem cell", "skeletal muscle satellite stem cell", "skeletal muscle satellite stem cell", "B cell", "endothelial cell", "macrophage")
cell_ontology_id <- c("CL:0000134", "CL:0008011", "CL:0008011", "CL:0000236", "CL:0000115", "CL:0000235")

tiss@meta.data[,'cell_ontology_class'] <- plyr::mapvalues(x = tiss@ident, from = cluster.ids, to = cell_ontology_class)
tiss@meta.data[,'cell_ontology_id'] <- plyr::mapvalues(x = tiss@ident, from = cluster.ids, to = cell_ontology_id)

tiss@meta.data[tiss@cell.names,'cell_ontology_class'] <- as.character(tiss@meta.data$cell_ontology_class)
tiss@meta.data[tiss@cell.names,'cell_ontology_id'] <- as.character(tiss@meta.data$cell_ontology_id)


TSNEPlot(object = tiss, do.label = TRUE, pt.size = 0.5, group.by='cell_ontology_class')
```


```{r}
tiss = BuildClusterTree(tiss)
```

## TODO run after this

```{r}
# Get markers for a particular cluster
cluster_markers = filter(tiss.markers, cluster == 3)$gene

DotPlot(tiss, genes.plot = cluster_markers[1:12], plot.legend = T)

```

We can repeat the above analysis on a subset of genes, defined using cluster IDs or expression or some other metadata. This is a good way to drill down and find substructure.

```{r}
# To subset data based on cell_ontology_class or other metadata, you can explicitly pass cell names
anno = 'B cell'
cells.to.use = tiss@cell.names[which(tiss@meta.data$cell_ontology_class == anno)]
subtiss <- SubsetData(object = tiss, cells.use = cells.to.use, do.center = F, do.scale = F)

subtiss <- NormalizeData(object = subtiss)
subtiss <- ScaleData(object = subtiss, vars.to.regress = c("nReads", "percent.ribo","Rn45s"))

#subtiss@scale.data = subtiss@data
#subtiss <- ScaleData(object = subtiss)
```



```{r}
subtiss <- FindVariableGenes(object = subtiss, do.plot = TRUE, x.high.cutoff = Inf, y.cutoff = 0.5)
subtiss <- RunPCA(object = subtiss, pcs.compute = 20)
print("de2")
subtiss <- ProjectPCA(object = subtiss, do.print = FALSE)
```


```{r}
PCElbowPlot(object = subtiss)
```

```{r, fig.height=6, fig.width=8}
PCHeatmap(object = subtiss, pc.use = 1:9, cells.use = 80, do.balanced = TRUE, label.columns = FALSE, num.genes = 18)
```


```{r}
subtiss <- FindClusters(object = subtiss, reduction.type = "pca", dims.use = 1:5, 
    resolution = 0.5, print.output = 0, save.SNN = TRUE, force.recalc = TRUE)
```


```{r}
subtiss <- RunTSNE(object = subtiss, dims.use = 1:10, seed.use = 10, check_duplicates = F, perplexity=25)
```

```{r}
# note that you can set do.label=T to help label individual clusters
TSNEPlot(object = subtiss, do.label = T)
```

Check expression of genes of interset.
```{r}
genes_to_check = c('Pax7', 'Pax3', 'Myod1', 'Myf5')

FeaturePlot(subtiss, genes_to_check, pt.size = 1)
```


```{r}
GenePlot(subtiss, 'Pax7', 'Pax3', use.raw = T)
```



When you save the annotated tissue, please give it a name.

```{r}
filename = here('00_data_ingest', '04_tissue_robj_generated', 
                     paste0("facs", tissue_of_interest, "_seurat_tiss.Robj"))
print(filename)
save(tiss, file=filename)
```


# Export the final metadata

So that Biohub can easily combine all your cell_ontology_classs, please export them as a simple csv.

```{r}
head(tiss@meta.data)
```


```{r}
filename = here('00_data_ingest', '03_tissue_cell_ontology_class_csv', 
                     paste0(tissue_of_interest, "_cell_ontology_class.csv"))
write.csv(tiss@meta.data[,c('plate.barcode','cell_ontology_class','cell_ontology_id')], file=filename)
```

