---
title: "Lung FACS Notebook"
output:
  html_document: default
  html_notebook: default
---

Specify the tissue of interest, run the boilerplate code which sets up the functions and environment, load the tissue object.

```{r}
tissue_of_interest = "Lung"
library(here)
source(here("00_data_ingest", "02_tissue_analysis_rmd", "boilerplate.R"))
load_tissue_facs(tissue_of_interest)
```

Visualize top genes in principal components

```{r, echo=FALSE, fig.height=4, fig.width=8}
PCHeatmap(object = tiss, pc.use = 18:30, cells.use = 500, do.balanced = TRUE, label.columns = FALSE, num.genes = 8)
```

```{r}
PCElbowPlot(object = tiss)
```


```{r}
n.pcs = 20
tiss <- FindClusters(object = tiss, reduction.type = "pca", dims.use = 1:n.pcs, resolution = 3, print.output = 0, save.SNN = TRUE, force.recalc = TRUE)
```


```{r}
tiss <- RunTSNE(object = tiss, dims.use = 1:n.pcs, seed.use = 10, check_duplicates = F)
```

```{r}
# note that you can set do.label=T to help label individual clusters
TSNEPlot(object = tiss, do.label = T, no.axes=T, pt.size = 1)
```

Check expression of genes of interset.
```{r}
genes_to_check = c('Pecam1', 'Epcam', 'Col1a1', 'Ptprc')

FeaturePlot(tiss, genes_to_check, pt.size = 1, cols.use = c('grey', 'red'))
```

```{r}
genes_to_check = c('Calca', 'Resp18', 'Ager', 'Scgb1a1')

FeaturePlot(tiss, genes_to_check, pt.size = 1, cols.use = c('grey', 'red'))
```

```{r}
genes_to_check = c('Foxj1', 'Cd79a', 'Klrb1c', 'Cd3e')

FeaturePlot(tiss, genes_to_check, pt.size = 1, cols.use = c('grey', 'red'))
```

```{r}
genes_to_check = c('H2-Aa', 'Lgals3', 'Itgal', 'Csf1r', 'Msln', 'Chad')

FeaturePlot(tiss, genes_to_check, pt.size = 1, cols.use = c('grey', 'red'))
```


Color by metadata, like plate barcode, to check for batch effects.
```{r}
TSNEPlot(object = tiss, do.return = TRUE, group.by = "plate.barcode")
```

Find differentially expressed markers.
```{r}
tiss.markers <- FindAllMarkers(object = tiss, only.pos = TRUE, min.pct = 0.25, 
    thresh.use = 0.25)
```

Display top 4 markers per cluster.
```{r}
tiss.markers %>% group_by(cluster) %>% top_n(4, avg_diff)
```


```{r}
# To change the y-axis to show raw counts, add use.raw = T.
VlnPlot(tiss, genes_to_check)
```


```{r}
table(tiss@ident)
```


```{r}
tiss = BuildClusterTree(tiss)
```

Get markers for a particular cluster
```{r}
cluster_markers = filter(tiss.markers, cluster == 4)$gene

DotPlot(tiss, genes.plot = cluster_markers[1:10])
```


## Assigning cell type identity to clusters

At a coarse level, we can use canonical markers to match the unbiased clustering to known cell types:

```{r}
# stash current cluster IDs
tiss <- StashIdent(object = tiss, save.name = "cluster.ids")

# enumerate current cluster IDs and the labels for them
cluster.ids <- c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20)
cell_ontology_class <-
  c(
  "Endothelial Cell Type I",
  "Stromal Cell Type I",
  "Endothelial Cell Type I",
  "Stromal Cell Type II &\n Mesothelial Cells",
  "Endothelial Cell Type I",
  "Aveolar Epithelial Cell Type II",
  "Natural Killer & T Cells",
  "Endothelial Cell Type II",
  "Immune Cell Type I",
  "Endothelial Cell Type III",
  'Immune Cell Type II',
  'Endothelial Cell Type IV',
  'Immune Cell Type III',
  'B cell',
  'Stromal Cell Type III',
  'Endothelial Cell Type II',
  'Stromal Cell Type IV',
  'Endothelial Cell Type IV',
  'Ciliated, Club, Neuroendocine,\n and Alveolar Epithelial Type I Cells',
  'Immune Cell Type IV',
  'Stromal Cell Type V'
  )

tiss@meta.data[,'free_annotation'] <- NA
tiss@meta.data[,'cell_ontology_class'] <- plyr::mapvalues(x = tiss@meta.data$cluster.ids, from = cluster.ids, to = cell_ontology_class)

TSNEPlot(object = tiss, do.label = TRUE, pt.size = 0.5, group.by='cell_ontology_class', no.legend = TRUE, no.axes = TRUE)
```


## Checking for batch effects


```{r}
TSNEPlot(object = tiss, do.return = TRUE, group.by = "mouse.sex")
```

Print a table showing the count of cells in each identity category from each plate.

```{r}
table(as.character(tiss@ident), as.character(tiss@meta.data$plate.barcode))
```


## Subset and iterate

We can repeat the above analysis on a subset of genes, defined using cluster IDs or expression or some other metadata. This is a good way to drill down and find substructure.

```{r}
# Get subtiss, eg, the endocrine clusters
subtiss <- SubsetData(object = tiss, ident.use = c(0, 1, 12),
  do.center = F, do.scale = F)

subtiss <- NormalizeData(object = subtiss)
subtiss@scale.data = subtiss@data
#subtiss <- ScaleData(object = subtiss)
#subtiss <- ScaleData(object = subtiss, vars.to.regress = c("nReads", "percent.ribo","Rn45s"))
```



```{r}
sub.n.pcs = 20
subtiss <- FindVariableGenes(object = subtiss, do.plot = TRUE, x.high.cutoff = Inf, y.cutoff = 0.5)
subtiss <- RunPCA(object = subtiss, pcs.compute = sub.n.pcs)

subtiss <- ProjectPCA(object = subtiss, do.print = FALSE)
```


```{r}
PCElbowPlot(object = subtiss)
```

```{r}
PCHeatmap(object = subtiss, pc.use = 1:6, cells.use = 500, do.balanced = TRUE, label.columns = FALSE, num.genes = 12)
```


```{r}
subtiss <- FindClusters(object = subtiss, reduction.type = "pca", dims.use = 1:sub.n.pcs, 
    resolution = 1.0, print.output = 0, save.SNN = TRUE, force.recalc = T)
```


```{r}
subtiss <- RunTSNE(object = subtiss, dims.use = 1:sub.n.pcs, seed.use = 10, check_duplicates = F)
```

```{r}
# note that you can set do.label=T to help label individual clusters
TSNEPlot(object = subtiss, do.label = T)
```


```{r}
write.csv(tiss@meta.data[,c('cell_ontology_class')],file =paste0(save_dir,"/", "SmartSeq2_", tissue_of_interest,"_cell_ontology_class.csv"))
```
