---
title: "R Notebook"
output: html_notebook
---

```{r}
library(Seurat)
library(cowplot)
library(data.table)
library(dplyr)
library(tidyr)
library(parallel)
library(cba)

tabula.dir <- "../../../tabula-muris/"
```

# Load Seurat object of data
```{r}
load("../../data/All_seurat_tiss.Robj")
```

# Load metadata and add to Seurat object
Loads Seurat object and list of TFs
Makes "annotation.2" metadata field which is tissue__cell_ontology_class. This is the field used to separate cells into classes for finding markers and doing correlations.
```{r}
metadata <- read.csv('../TM_facs_metadata.csv')
metadata <- metadata %>%  filter(cell_ontology_class!='unknown' & !is.na(tissue) & !is.na(cell_ontology_class))
metadata$tissue <- make.names(metadata$tissue)
metadata$cell_ontology_class <- make.names(metadata$cell_ontology_class)

# make "annotation.2"
metadata  <- metadata %>% mutate(annotation.2 = paste0(tissue, "__", cell_ontology_class))

tissue_colors <- read.csv(file.path(tabula.dir,'00_data_ingest/15_color_palette/tissue_colors.csv'))
colnames(tissue_colors) <- c('tissue','tiss.color')
tissue_colors$tissue <- make.names(tissue_colors$tissue)

metadata <- merge(metadata, tissue_colors, by = 'tissue')

rownames(metadata) <- metadata$cell # need this to add to Seurat object
metadata <- metadata[tiss@cell.names, ]

# Add metadata
tiss <- AddMetaData(tiss, metadata)
sum(is.na(tiss@meta.data$annotation.2))
length(unique(tiss@meta.data$annotation.2))
tiss <- SetAllIdent(tiss, 'annotation.2')

# Make gene names R compatible
rownames(tiss@data) <- make.names(rownames(tiss@data))
rownames(tiss@raw.data) <- make.names(rownames(tiss@raw.data))
rownames(tiss@scale.data) <- make.names(rownames(tiss@scale.data))
gc()
```

# Load TFs 
```{r}
# (1140 genes from MGI, filtered by GO Term = "DNA binding transcription factor activity", GO ID ?)
tfs      <- fread('../GO_term_summary_20171110_222852.csv')
tf.names <- as.character( tfs %>% distinct(Symbol) %>% pull(Symbol) )
tf.names <- make.names(tf.names)
tf.names <- tf.names[tf.names %in% rownames(tiss@data)]
length(tf.names)

# remove genes with very low expression in data
x=apply(as.matrix(tiss@data[tf.names, ]) > 0, 1, sum) 
genes.expr <- names(x[x > 5])
tf.names <- tf.names[tf.names %in% genes.expr]

# Remove IEGs
iegs <- as.character(read.csv(file.path(tabula.dir,"00_data_ingest/20_dissociation_genes/genes_affected_by_dissociation_unix.csv"), header = T)[,1])
tf.names <- tf.names[!tf.names %in% iegs]
length(tf.names)
```

# Subsample data
```{r}
tiss.subsamp <- SubsetData(tiss, max.cells.per.ident = 60, subset.raw = T)
head(unique(tiss.subsamp@meta.data$annotation.2))
```

# Set identity to 

# Function to calculate per-cell type expression averages
```{r}
group_averages <- function(mat, groups){
  group_names = unique(groups)
  means = matrix(0, dim(mat)[1], length(group_names))
  colnames(means) = group_names
  rownames(means) = rownames(mat)
  for(group in group_names){
    means[,group] = Matrix::rowMeans(mat[,groups == group,drop=FALSE])
  }
  means
}

# annotation.2.means <- group_averages(tiss.subsamp@data, tiss@ident)

```

# Correlation + plotting functions
Revision changes: correlations are calculated on the celltype/tissue group averages (arithmetic mean of log1p values),
rather than on single cells from subsampled dataset as in the initial submission.
```{r}
# cluster genes based on expression correlation and plot correlogram
# returns output of barb.cormap, which includes the list of genes in the order they appear in the correlogram
plot.correlogram <- function(data, pval_cutoff, avg_logFC_cutoff, nmarkers.per.celltype, markerfile="TF_markers.csv",height=10){
  
  # read in markers
  markers <- read.csv(markerfile, row.names = 1)
  markers$gene <- rownames(markers)
  markers$avg_logFC <- as.numeric(as.character(markers$avg_logFC))
  markers$p_val <- as.numeric(as.character(markers$p_val))
  
  print(head(markers))
  
  # Get top genes by cell type
  genes.use <- as.character(unique(markers %>% dplyr::group_by(cluster) %>% 
                      filter(gene %in% tf.names, p_val < pval_cutoff, avg_logFC > avg_logFC_cutoff) %>% 
                      dplyr::top_n(n=nmarkers.per.celltype, wt=avg_logFC) %>% 
                      pull(gene)))

  print(paste("genes plotted: ",length(genes.use)))
  # Calculate correlations
  # mat.use <- t(as.matrix(tiss@data[genes.use, ]))
  
  # use cell type averages instead (changed for revision)
  annotation.2.means <- group_averages(data@data, data@ident)
  mat.use <- t(annotation.2.means[genes.use, ])
  tf.cor    <- cor(mat.use)
  print(colnames(tf.cor))
  enrich.score <- dcast(markers, gene ~ cluster, value.var = 'avg_logFC')
  rownames(enrich.score) <- enrich.score[,"gene"]
  enrich.score <- enrich.score[, 2:ncol(enrich.score)]
  # change rownames to the cell type and tissue that each gene is enriched in (ordering by avg_logFC above)
  topIDenriched   <- sapply(colnames(tf.cor), function(x) {
    names(sort(t(enrich.score)[, x],  decreasing = T))[1]}) 
  colnames(tf.cor) <- topIDenriched
  
  correlo.out <- barb.cormap(tf.cor,  'TF_cormap.pdf', height =height, width = height)
  
  # Generate row colors of heatmap corresponding to highest-expressing cell type
  topID.plotorder <- correlo.out[[2]]
  meta.summary <- data@meta.data %>% distinct(tissue, cell_ontology_class, tiss.color, annotation.2)
  print(head(meta.summary))
  ntypes=length(unique(meta.summary$cell_ontology_class))
  tmp <- colorRampPalette(brewer.pal(min(ntypes, 11), 'Paired'))(ntypes)
  annot_colors <- data.frame(annot.colors = tmp, cell_ontology_class = unique(meta.summary$cell_ontology_class))
  write.csv(annot_colors, file=paste0('annotColors.csv'))
  
  print(head(annot_colors))
  meta.summary <- merge(meta.summary, annot_colors, by = 'cell_ontology_class')
  print(head(meta.summary))

  plot.colors <- data.frame(annotation.2 = topID.plotorder)
  print(head(plot.colors))

  plot.colors <- merge(plot.colors, meta.summary %>% select(annotation.2, annot.colors, tiss.color), by = 'annotation.2')
  print((plot.colors))
  plot.colors <- plot.colors[match(topID.plotorder, plot.colors$annotation.2), ]
  print((plot.colors))
  plot.colors$ymin <- 0.1*(0:(nrow(plot.colors)-1))
  plot.colors$ymax <- 0.1*(1:(nrow(plot.colors)))
  print((plot.colors))
  plot.colors$tiss.color <- toupper(plot.colors$tiss.color)
    plot.colors$annot.colors <- toupper(plot.colors$annot.colors)


  require(grDevices)
  pdf( 'TF_cormap_rowcolors_AnnotRight_TissueLeft.pdf', height = 10, width = 4)
    plot(c(0, 2), c(0, max(plot.colors$ymax) + 1), type = "n", xlab = "", ylab = "",
         main = "plot colors")
    rect(0,plot.colors$ymin, 1 , plot.colors$ymax, col = plot.colors$annot.colors, border = NA)
    rect(1,plot.colors$ymin, 2 , plot.colors$ymax, col = plot.colors$tiss.color, border = NA)
  dev.off()

  return(correlo.out)
}

barb.cormap <- function(mat.cor,fname,width=12,height=12,method="complete",cex=0.5,mincor=-1,maxcor=1){
  require(lattice)
  require(cba)
  rowdist <- dist(mat.cor)
  coldist <- dist(mat.cor, by_rows = F)
  hc.cor <- hclust(coldist, method=method)
  hr.cor <- hclust(rowdist, method=method)

  optimal.row <- order.optimal(rowdist,hr.cor$merge)
  optimal.col <- order.optimal(coldist,hc.cor$merge)
 
  ord.row <- optimal.row$order
  ord.col <- optimal.col$order
  
  plt = levelplot(mat.cor[ord.row,ord.col],xlab=NULL,ylab=NULL,
                  at=do.breaks(c(mincor-0.01,maxcor+0.01),19),scales=list(x=list(rot=90),cex=cex),
                  colorkey=list(space="top"),
                  col.regions=colorRampPalette(c("dodgerblue4", "dodgerblue", "white", "lightcoral", "firebrick4"), space="Lab"))
  pdf(fname,width=width,height=height)
    print(plt)
  dev.off()
  
  return(list(rownames(mat.cor[ord.row, ]), colnames(mat.cor[, ord.col]), plt, hc.cor, hr.cor))
}

```

# All cell types TF
Make a list of TFs with cell type specificity to use for correlation analysis.
Version 3 (preparing for final revision, 5/16/18): Seurat FindAllMarkers is faster. FindMarkers now uses much more memory and cannot be easily parallelized w/o crashing (even w/64GB RAM). FindAllMarkers uses 18GB RAM. -->
```{r}
res=FindAllMarkers(tiss.subsamp, genes.use = tf.names,
                   only.pos = T, test.use = "wilcox", min.diff.pct = 0.1, print.bar = F, do.print = T)
write.csv(res, "TF_markers.csv")
```
# compute and plot correlations
```{r}
correlo.all <- plot.correlogram(data = tiss.subsamp, pval_cutoff = 10^-3.5, avg_logFC_cutoff = 0.15, nmarkers.per.celltype = 6, height = 30)
```


