---
title: "Cluster Compactness"
output: html_notebook
---

```{r}
library(dplyr)
library(Seurat)
library(here)
load(here('00_data_ingest','11_global_robj','FACS_all.Robj'))
```

## Compactness

We can measure the heterogeneity of a cluster by the average distance between pairs of its members.
We can measure the 'mixing' of groups within a population (eg cell types w/in a cluster) by the average distance between pairs of points in each population.

```{r}
avg_pairwise_distance <- function(mat, n_samples = 1000){
  n = dim(mat)[1]
  total_distance = 0
  for(s in 1:n_samples){
    samples = sample(n,2)
    diff = mat[samples[1],] - mat[samples[2],]
    distance = sqrt(sum(diff**2))
    total_distance = total_distance + distance
  }
  average_distance = total_distance/n_samples
  average_distance
}

# test on identity matrix, where all pairwise distances are sqrt(2)
test_mat = matrix(c(1,0,0,0,1,0,0,0,1), nrow = 3)
avg_pairwise_distance(test_mat)

# n.pcs
avg_cluster_distance <- function(j, n.pcs = 100){
  cells = tiss_FACS@meta.data %>% filter(cluster == j) %>% pull(cell)
  
  mat = tiss_FACS@dr$pca@cell.embeddings[cells, 1:n.pcs]
  avg_pairwise_distance(mat)
}

avg_pairwise_distance2 <- function(mat1, mat2, n_samples = 1000){
  n1 = dim(mat1)[1]
  n2 = dim(mat2)[1]
  total_distance = 0
  for(s in 1:n_samples){
    sample1 = sample(n1, 1)
    sample2 = sample(n2, 1)
    diff = mat1[sample1,] - mat2[sample2,]
    distance = sqrt(sum(diff**2))
    total_distance = total_distance + distance
  }
  average_distance = total_distance/n_samples
  average_distance
}

test_mat_1 = matrix(c(1, 1.1, 0, 0.1), ncol = 2)
test_mat_2 = matrix(c(0, 0.1, 1, 1.1), ncol = 2)
avg_pairwise_distance2(test_mat_1, test_mat_2)

avg_pairwise_distance_set <- function(mat, ids, n_samples = 1000, min_id_size = 1, verbose = F){
  id_list = unique(ids)
  # Keep only those ids occuring at least twice
  id_list = id_list[sapply(id_list, function(x){
      if(sum(ids == x) < min_id_size){
        if(verbose){
        print(paste0("Too few instances of ", x))
        }
        FALSE
      }
      else{
        TRUE
      }
    }
    )]
  
  n_ids = length(id_list)

  distance_mat = matrix(NA, nrow = n_ids, ncol = n_ids)
  rownames(distance_mat) = id_list
  colnames(distance_mat) = id_list
  for(i in 1:length(id_list)){
    id_1 = id_list[i]
    # First do the self-distances
    mat_1 = mat[ids == id_1,]
    distance_mat[i,i] = avg_pairwise_distance(mat[ids == id_1,])
    if(i < length(id_list)){
      for(j in (i+1):length(id_list)){
        id_2 = id_list[j]
        mat_2 = mat[ids == id_2,]
        distance_mat[i,j] = avg_pairwise_distance2(mat_1, mat_2)
        distance_mat[j,i] = distance_mat[i,j]
      }
    }
  }
  distance_mat
}

avg_pairwise_distance_set(rbind(test_mat_1, test_mat_2), c('a','a','b','b'), min_id_size = 2)

average_cluster_differences_set <- function(j, n.pcs = 100, 
                                            min_id_size = 5, dropna = T){
  cluster_metadata = tiss_FACS@meta.data %>% filter(cluster == j & (!(dropna & is.na(cell_ontology_class))))
  
  cells = cluster_metadata %>% pull(cell)
  
  mat = tiss_FACS@dr$pca@cell.embeddings[cells, 1:n.pcs]

  class_sizes = cluster_metadata %>% 
    mutate(cell_ontology_class = 
             ifelse(is.na(cell_ontology_class), "NA", cell_ontology_class)) %>%
    mutate(class = paste0(cell_ontology_class, " (", tissue, ")")) %>% group_by(class) %>%
    count() %>% mutate(class_with_size = paste0(class, " (", n,")"))
  
  
  classes = cluster_metadata %>% 
    mutate(cell_ontology_class = 
             ifelse(is.na(cell_ontology_class), "NA", cell_ontology_class)) %>%
    mutate(class = paste0(cell_ontology_class, " (", tissue, ")")) %>% 
    pull(class)
  
  classes = plyr::mapvalues(x = classes, 
                            from = class_sizes$class,
                            to = class_sizes$class_with_size)
  
  avg_pairwise_distance_set(mat, classes, min_id_size = min_id_size)
}
```


```{r}
library(gplots)
library(viridis)

plot_dist <- function(dists, relative = TRUE, xmin = 10, xmax = 40){
  ncol = 100
  n_types = dim(dists)[1]
  if(relative){
    heatmap.2(dists, trace = "none", margins=c(50,50), col = viridis(ncol), 
            symm = TRUE, revC = TRUE, breaks = seq(xmin, xmax, (xmax - xmin)/ncol), dendrogram = "none", density.info = "none")
  }
  else{
    heatmap.2(dists, trace = "none", margins=c(50,50), col = viridis(ncol), 
            symm = TRUE, revC = TRUE)
  }
}
```

Compute the average distance between all pairs of points.

```{r}
global_avg = avg_pairwise_distance(tiss_FACS@dr$pca@cell.embeddings[, 1:100], n_samples = 10000)
```

Produce all plots.

```{r}
for (i in 0:53){
  dists = average_cluster_differences_set(i, min_id_size = 5)
  if(dim(dists)[1] > 1){
    pdf(here('32_cluster_heterogeneity', 'figures', paste0("cluster_breakdown_", i, ".pdf")), width = 20, height = 20)
    plot_dist(dists, xmax = global_avg)
    dev.off()
  }
}
```

## Statistics for manuscript

Cells in cluster 3, and breakdown.
```{r}
tiss_FACS@meta.data %>% filter(cluster == 3) %>% count()
tiss_FACS@meta.data %>% filter(cluster == 3) %>% group_by(cell_ontology_class)  %>% count() %>% filter(n > 5)
```

The Thymus cells in cluster 3 express _Pecam1_.

```{r}
FetchData(tiss_FACS, c('cell_ontology_class', 'cluster', 'Pecam1', 'tissue')) %>% 
  filter(cluster == 3 & tissue == 'Thymus')
```

```{r}
distances = sapply(0:53, function(i) {average_cluster_differences_set(i, min_id_size = 5)}, simplify = F)
```

```{r}
geo_normalize <- function(dists){
  dists / (sqrt(diag(dists)) %o% sqrt(diag(dists)))
}
normalized_distances = lapply(distances, geo_normalize)
```

We discuss 2,3,48,24 in the text.

```{r, fig.width = 30, fig.height = 30}
cluster = 2
plot_dist(normalized_distances[[cluster + 1]], relative = T, xmax = 3, xmin = 1)
plot_dist(distances[[cluster + 1]], relative = T, xmax = 40)
```

```{r}
positive_examples = c(1, 2, 3, 48, 24)
negative_examples = c(31, 32, 14, 53, 35, 9, 21, 41)
```

Add filter to include only things with multiple tissues...that leaves just 25 clusters.

```{r}
tissue_count = tiss_FACS@meta.data %>% group_by(cluster, tissue) %>% count() %>% filter(n >= 5) %>% ungroup() %>%
  group_by(cluster) %>% count() %>% rename(tissues = nn) %>% ungroup() %>% mutate(cluster = as.character(cluster))
tissue_count
```

Distributions of heterogeneity scores.

```{r}
max_dist = sapply(distances, max)/global_avg
avg_dist = sapply(distances, mean)
worst_dist = sapply(normalized_distances, max)
df = tibble(cluster = as.character(0:53),
            max_dist = max_dist,
            worst_dist = worst_dist, 
            n_types = sapply(distances, function(x){dim(x)[1]}),
            rank = rank(max_dist)
            )
df = df %>% left_join(tissue_count, by = 'cluster')
```

```{r}
df %>% filter(tissues > 1) %>% ggplot(aes(fct_reorder(cluster, rank), max_dist)) + geom_bar(stat = 'identity') +
   geom_hline(aes(yintercept = 1.0)) + ylab('heterogeneity score') + xlab('cluster')
ggsave(here('32_cluster_heterogeneity', 'figures', 'heterogeneity_score.pdf'))
```


