---
title: "Supplementary Figure 1"
output: html_document
---

```{r}
library(tidyverse)
library(stringr)
library(Seurat)
```

```{r}
df_all <- tibble(cell = character(), tissue = character(), subtissue = character(), nGene = integer(), method = character())
```

# FACS data

```{r}
load(file=here("00_data_ingest", "11_global_robj", "FACS_all.Robj"))
```

```{r}
df_FACS <- as_tibble(tiss_FACS@meta.data[c('tissue', 'subtissue', 'nGene')])
df_FACS <- rownames_to_column(df_FACS, "cell")
df_FACS['method'] <- 'FACS'
```


# Droplet data

```{r}
load(file=here("00_data_ingest", "11_global_robj", "droplet_all.Robj"))
```

```{r}
df_droplet <- as_tibble(tiss_droplet@meta.data[c('tissue', 'subtissue', 'nGene')])
df_droplet <- rownames_to_column(df_droplet, "cell")
df_droplet['method'] <- 'droplet'
```

# Han data

Load the processed and annotated data.

```{r}
han_metadata <- read_csv(here('00_data_ingest','han_data','han_metadata.csv'))
han_metadata_comparable = han_metadata %>% filter(!is.na(tissue)) %>% filter(is.na(subtissue))
han_metadata_comparable
```

```{r}
df_microwell <- tibble(cell = character(), tissue = character(), subtissue = character(), nGene = integer(), method = character())

for(i in 1:nrow(han_metadata_comparable)){
  print(paste0("Loading ", han_metadata_comparable$filename[i]))
  raw.data <- read.table(here('00_data_ingest','han_data',han_metadata_comparable$filename[i]), header = TRUE)
  df <- tibble(cell = colnames(raw.data))
  df['tissue'] <- han_metadata_comparable$tissue[i]
  df['subtissue'] <- han_metadata_comparable$subtissue[i]
  df['method'] <- 'microwell'
  df['nGene'] <- Matrix::colSums(raw.data > 0)
  for(j in 0:4){
    df[paste0('nGene.', j)] <- Matrix::colSums(raw.data > j)
  }
  df_microwell <- bind_rows(df_microwell, df)
}
```

# Merge and Plot

```{r}
df_all = bind_rows(df_FACS, df_droplet, df_microwell)
```

```{r}
df_all %>% group_by(tissue, method) %>% count()
```


Select the tissues processed using all methods.

```{r}
common_tissues <- as.character(df_all %>% 
                                 group_by(tissue, method) %>% summarize(count = n()) %>% 
                                 ungroup() %>% group_by(tissue) %>% summarize(count = n()) %>% 
                                 filter(count > 1) %>% pull(tissue))
common_tissues
```

Plot the histograms of the number of genes per cell for each tissue and method.

```{r, fig.width = 12, fig.height = 8}
df_all %>% filter(nGene > 500) %>% 
filter(tissue %in% common_tissues) %>% ggplot(aes(nGene, ..density..,colour = method)) + 
  geom_freqpoly(binwidth = 100) +
  facet_wrap(~ tissue, ncol = 4) + 
  ggtitle("Number of genes expressed")

#ggsave(here("11_supplementary_figure1","sfigure1.pdf"), width = 7, height = 7, units = "in")
```

Boxplots of the same distributions.

```{r, fig.width = 12, fig.height = 8}
df_all %>% filter(nGene > 500) %>% 
filter(tissue %in% common_tissues) %>% ggplot(aes(x=tissue, y=nGene, fill=method)) + 
    geom_boxplot() + 
  ggtitle("Number of genes expressed")

ggsave(here("11_supplementary_figure1","sfigure1.pdf"), width = 14, height = 7, units = "in")
```

## Dependency of boxplots on cutoff

For UMI we do cutoffs of 0,1,2,3,4. For reads we do 10 times those numbers, as 10 is roughly the average reads/UMI, though it varies vastly by eg 10x run.

```{r}
for(i in c(0,1,2,3,4)){
  df_FACS[paste0("nGene.",i)] = Matrix::colSums(tiss_FACS@raw.data[,df_FACS %>% pull(cell)] > 10*i)
}

for(i in c(0,1,2,3,4)){
  df_droplet[paste0("nGene.",i)] = Matrix::colSums(tiss_droplet@raw.data[,df_droplet %>% pull(cell)] > i)
}
```


```{r, fig.width = 12, fig.height = 12}

df_FACS %>% filter(nGene > 500) %>% mutate(percentnGene.0 = nGene.0/nGene,
                                              percentnGene.1 = nGene.1/nGene,
                                              percentnGene.2 = nGene.2/nGene,
                                              percentnGene.3 = nGene.3/nGene,
                                              percentnGene.4 = nGene.4/nGene) %>%
  gather(percentnGene.1:percentnGene.4, key=cutoff, value = percentnGene.cutoff) %>%
  separate(cutoff, c("std","threshold")) %>% select(-std) %>% mutate(Read.threshold = as.character(10*as.integer(threshold))) %>%
  ggplot(aes(x=tissue, y=percentnGene.cutoff, fill=Read.threshold)) +
  geom_boxplot() +
  xlab("Tissue") + ylab("Fraction of detected genes present above threshold") + ylim(0,1) +
  ggtitle("FACS") + 
  coord_flip()

ggsave(here("11_supplementary_figure1","sfigure1-facs.pdf"), width = 7, height = 10, units = "in")

df_droplet %>% filter(nGene > 500) %>% mutate(percentnGene.0 = nGene.0/nGene,
                                              percentnGene.1 = nGene.1/nGene,
                                              percentnGene.2 = nGene.2/nGene,
                                              percentnGene.3 = nGene.3/nGene,
                                              percentnGene.4 = nGene.4/nGene) %>%
  gather(percentnGene.1:percentnGene.4, key=cutoff, value = percentnGene.cutoff) %>%
  separate(cutoff, c("std","UMI.threshold")) %>% select(-std) %>%
  ggplot(aes(x=tissue, y=percentnGene.cutoff, fill=UMI.threshold)) +
  geom_boxplot() +
  xlab("Tissue") + ylab("Fraction of detected genes present above threshold") + ylim(0,1) +
  ggtitle("Droplet") + 
  coord_flip()
ggsave(here("11_supplementary_figure1","sfigure1-droplet.pdf"), width = 7, height = 10, units = "in")


df_microwell %>% filter(nGene > 500) %>% mutate(percentnGene.0 = nGene.0/nGene,
                                              percentnGene.1 = nGene.1/nGene,
                                              percentnGene.2 = nGene.2/nGene,
                                              percentnGene.3 = nGene.3/nGene,
                                              percentnGene.4 = nGene.4/nGene) %>%
  gather(percentnGene.1:percentnGene.4, key=cutoff, value = percentnGene.cutoff) %>%
  separate(cutoff, c("std","UMI.threshold")) %>% select(-std) %>%
  ggplot(aes(x=tissue, y=percentnGene.cutoff, fill=UMI.threshold)) +
  geom_boxplot() +
  xlab("Tissue") + ylab("Fraction of detected genes present above threshold") + ylim(0,1) +
  ggtitle("Microwell") + 
  coord_flip()
ggsave(here("11_supplementary_figure1","sfigure1-microwell.pdf"), width = 7, height = 10, units = "in")

```

Median number of genes for each tissue, method.

```{r}
df_all %>% group_by(tissue, method) %>% summarize(med = median(nGene))
```

