---
title: "R Notebook"
output: html_notebook
---


```{r}
rm(list=ls())
library(Seurat)
library(cowplot)
library(data.table)
library(dplyr)
library(tidyr)
library(parallel)
library(cba)
library(doMC)
library(RColorBrewer)

tabula.dir <- "../../../tabula-muris/"
```

# Load Seurat object of data
```{r}
load(file.path(tabula.dir,"data/All_seurat_tiss.Robj"))
```

# Load metadata and add to Seurat object
Loads Seurat object and list of TFs
Makes "annotation.2" metadata field which is tissue__cell_ontology_class. This is the field used to separate cells into classes for finding markers and doing correlations.
```{r}
metadata <- read.csv('../TM_facs_metadata.csv')
metadata <- metadata %>%  filter(cell_ontology_class!='unknown' & !is.na(tissue) & !is.na(cell_ontology_class))
metadata$tissue <- make.names(metadata$tissue)
metadata$cell_ontology_class <- make.names(metadata$cell_ontology_class)

# make "annotation.2"
metadata  <- metadata %>% mutate(annotation.2 = paste0(tissue, "__", cell_ontology_class))

tissue_colors <- read.csv(file.path(tabula.dir,'00_data_ingest/15_color_palette/tissue_colors.csv'))
colnames(tissue_colors) <- c('tissue','tiss.color')
tissue_colors$tissue <- make.names(tissue_colors$tissue)

metadata <- merge(metadata, tissue_colors, by = 'tissue')

rownames(metadata) <- metadata$cell # need this to add to Seurat object
metadata <- metadata[tiss@cell.names, ]

# Add metadata
tiss <- AddMetaData(tiss, metadata)
sum(is.na(tiss@meta.data$annotation.2))
length(unique(tiss@meta.data$annotation.2))
tiss <- SetAllIdent(tiss, 'annotation.2')

# Make gene names R compatible
rownames(tiss@data) <- make.names(rownames(tiss@data))
rownames(tiss@raw.data) <- make.names(rownames(tiss@raw.data))
rownames(tiss@scale.data) <- make.names(rownames(tiss@scale.data))
gc()
```

# Load TFs 
```{r}
# (1140 genes from MGI, filtered by GO Term = "DNA binding transcription factor activity", GO ID ?)
tfs      <- fread('../GO_term_summary_20171110_222852.csv')
tf.names <- as.character( tfs %>% distinct(Symbol) %>% pull(Symbol) )
tf.names <- make.names(tf.names)
tf.names <- tf.names[tf.names %in% rownames(tiss@data)]
length(tf.names)

# remove genes with very low expression in data
x=apply(as.matrix(tiss@data[tf.names, ]) > 0, 1, sum) 
genes.expr <- names(x[x > 5])
tf.names <- tf.names[tf.names %in% genes.expr]

# Remove IEGs
iegs <- as.character(read.csv(file.path(tabula.dir,"00_data_ingest/20_dissociation_genes/genes_affected_by_dissociation_unix.csv"), header = T)[,1])
tf.names <- tf.names[!tf.names %in% iegs]
length(tf.names)
```

# Subsample data
```{r}
tiss.subsamp <- SubsetData(tiss, max.cells.per.ident = 60, subset.raw = T)
head(unique(tiss.subsamp@meta.data$annotation.2))
```

# Differential expression of TFs (FindAllMarkers)
Make a list of TFs with cell type specificity to use for correlation analysis.
```{r}
res=FindAllMarkers(tiss.subsamp, genes.use = tf.names,
                   only.pos = T, test.use = "wilcox", min.diff.pct = 0.1, print.bar = F, do.print = T)
write.csv(res, "TF_markers.csv")
```

# Best 4 combinatorial markers
Positive hit only if a cell expresses all 4 markers
```{r}
output <- data.frame(cell.type=character(0), 
                       ppv=numeric(0),
                       recall=numeric(0),
                       TF_1 = character(0), 
                       TF_2 = character(0), 
                       TF_3 = character(0),
                      TF_4 = character(0))

# Iterate through cell types, looking for the markers that best distinguish each one
celltypes <- as.character(unique(tiss.subsamp@ident))

for(target.type in celltypes) {
  markers <- read.csv('TF_markers.csv') # differential expression by Seurat, with subsampled cells
  genes.use <- as.character(markers %>% 
                        filter(cluster==target.type,
                               gene %in% tf.names, 
                               p_val < 1e-3, 
                               avg_logFC > .2,
                               pct.1-pct.2 > 0.1,
                               pct.1 > 0.3) %>% 
                          arrange(desc(avg_logFC)) %>%
                        pull(gene))[1:30]
  genes.use <- genes.use[!is.na(genes.use)]
  
  if(length(genes.use) > 4){  # Only look for TFs in a cell type if it has at least 4 enriched TFs that pass the above filters


    cells.1 = tiss.subsamp@cell.names[tiss.subsamp@ident == target.type]
    cells.2 = tiss.subsamp@cell.names[tiss.subsamp@ident != target.type]
    
    topmarkers.bin <- tiss.subsamp@data[genes.use, ] > 0.35
    topmarkers.1 <- topmarkers.bin[, cells.1]
    topmarkers.2 <- topmarkers.bin[, cells.2]
    
  
    # get sensitivity and specificity of genes in these top 50 (binary on/off expression)
    combo.4 <- combn(genes.use, 4)
    
    registerDoMC(detectCores())
    
    # greedy-2-1-1 protocol (try every combo of 2 genes, then test adding one gene, then test adding another)
    # then you can see for each cell type how many genes you really need
    tp.fp <- foreach(i = 1:ncol(combo.4),.combine=cbind) %dopar% {
      
      # positive: detect all 4 genes
      # negative: detect < 4 genes
      tp <- sum(Matrix::colSums(topmarkers.1[combo.4[, i], ]) > 3)
      fp <- sum(Matrix::colSums(topmarkers.2[combo.4[, i], ]) > 3)
      
      c(tp, fp)
    }
    tp.fp <- t(tp.fp)
    tp.fp <- as.data.frame(tp.fp)
    colnames(tp.fp) <- c('tp','fp')
    tp.fp$ppv <- tp.fp$tp / (tp.fp$tp + tp.fp$fp) # ppv is true pos / true pos + false pos
    tp.fp$recall <- tp.fp$tp / length(cells.1) # recall is true pos / all in target
    
    tp.fp$score <- tp.fp$ppv + 10*tp.fp$recall
    top.combo.ids <- order(tp.fp$score, decreasing = T)[1:3]
    combo.4[, top.combo.ids]
    tp.fp[top.combo.ids, ]
    
    output <- rbind(output,data.frame(cell.type=rep(target.type, 3), 
                         ppv=tp.fp[top.combo.ids, 'ppv'],
                         recall=tp.fp[top.combo.ids, 'recall'],
                         TF_1 = combo.4[1, top.combo.ids],
                         TF_2 = combo.4[2, top.combo.ids],
                         TF_3 = combo.4[3, top.combo.ids],
                         TF_4 = combo.4[4, top.combo.ids]))
    
  }
}


output$score <- output$ppv + output$recall
output %>% arrange(desc(score)) %>% group_by(cell.type) %>% slice
write.csv(output, file = 'best_4_markers.csv', row.names = F, quote=F)
```



## Bar plots of precision/recall
```{r}


top4markers <- read.csv('best_4_markers.csv')
top4markers$score <- top4markers$recall + 2*top4markers$ppv
top4.summary <- top4markers %>% group_by(cell.type) %>% filter(recall > 0.2) %>% slice(which.max(score)) %>% arrange(desc(score)) 

plotdir <- 'specificity_plots'
dir.create(plotdir)
plist=list()
for(i in 1:20){ # plot the barplots of the top 20 cell types by precision/recall of their best TF combo
  genecombo <- as.character(t(top4.summary[i, c("TF_1","TF_2","TF_3","TF_4")]))
  target.type <- as.character(t(top4.summary[i, "cell.type"]))
  cells.1 = tiss.subsamp@cell.names[tiss.subsamp@ident == target.type]
  cells.2 = tiss.subsamp@cell.names[tiss.subsamp@ident != target.type]
  
  topmarkers.bin <- tiss.subsamp@data[genecombo, ] > 0.35
  
  # number of cells expressing all 4 genes per subtype
  all.expr <- t(as.matrix(Matrix::colSums(topmarkers.bin) > 3))
  
  group_sums <- function(mat, groups){
    group_names = unique(groups)
    sums = matrix(0, dim(mat)[1], length(group_names))
    colnames(sums) = group_names
    rownames(sums) = rownames(mat)
    for(group in group_names){
      sums[,group] = sum(mat[,groups == group,drop=FALSE])
    }
    sums
  }
  df.ppv <- data.frame(t(group_sums(all.expr, tiss.subsamp@ident)))
  colnames(df.ppv) <- "num.cells"
  df.ppv$cell.type <- rownames(df.ppv)
  df.ppv <- df.ppv[df.ppv$num.cells > 0, ]
  df.ppv$cell.type <- factor(df.ppv$cell.type, levels=df.ppv$cell.type[order(-df.ppv$num.cells)])
  df.ppv$plt.group <- 'ppv'
  
  # calculate recall
  num.on <- df.ppv$num.cells[df.ppv$cell.type==target.type]
  num.off <- sum(tiss.subsamp@ident==target.type) - df.ppv$num.cells[df.ppv$cell.type==target.type]
  df.recall <- data.frame(cell.type = c(target.type, paste0("Neg_",target.type)), 
                          num.cells = c(num.on, num.off))
  df.recall$plt.group <- 'recall'
  
  
  df.plot <- rbind(df.ppv, df.recall)
  n.col <- length(unique(df.plot$cell.type))
  cols <- c(brewer.pal(9,'Set1')[1:(n.col-1)], 'grey60')
  p <- ggplot(df.plot, aes(plt.group, num.cells, fill = cell.type)) +
    geom_bar(stat = 'identity',position = position_stack(reverse = TRUE), width = .7) +
    scale_fill_manual(values=cols) +
    theme(axis.title = element_blank(), 
          axis.text.x = element_text(angle = 45, hjust = 1, size=16),
          legend.title = element_blank(),
          legend.position = "bottom",
          legend.direction='vertical',
          legend.text = element_text(size=16),
          axis.text.y = element_text(size=16), 
          plot.title = element_text(size=20, margin = margin(b=15), hjust = 0.5), 
          plot.margin = unit(c(0.5,1.3,0,1.3), 'cm')) +
    ggtitle(paste0(paste(genecombo[1:2], collapse = " / ")," /\n",
                   paste(genecombo[3:4], collapse = " / ")))
  
  
  legend <- get_legend(p)
  p3 <- plot_grid(p + theme(legend.position = 'none'), legend, nrow=2,rel_heights = c(.8,.4))
  
  
  save_plot(paste0(plotdir,"/", target.type, '.pdf'),p3, nrow=2, base_height = 3, base_aspect_ratio = 1.5)
  
}


```

## Number of cell types with a highly specific set of genes
```{r}


top4markers <- read.csv('best_4_markers.csv')
top4markers$score <- top4markers$recall + 2*top4markers$ppv
numSpecificTypes <- length(unique( top4markers %>% 
                                     group_by(cell.type) %>% 
                                     filter(recall > 0.3, ppv > 0.3) %>% 
                                     pull(cell.type)))

print(numSpecificTypes)

```


# RF method (not used)
In Aevermann et al  http://dx.doi.org/10.1101/257352, they develop a "necessary and sufficient" set of genes using the following protocol.

1. classification model is constructed from one hundred thousand trees using Information Gain Ratio
as the splitting criteria, where each decision tree is generated using the default bagging parameters
- the square root of the number of features and a bootstrap of samples equal to the training set size.
2.  list of genes can be ranked by their usefulness in distinguishing one cell type

3. In the second step, single decision trees are constructed using the
first gene from the ranked list, the first two genes, the first three genes, etc. 

4. We restrict each gene to being used in only one branch per tree, and find the
optimal classification for the target cluster only, rather than the overall classification score. The
addition of genes from the ranked list is stopped when an optimal classification or stable tree
topology is achieved. 

5. The minimum number of genes used to produce this optimal result corresponds to the set of necessary and sufficient marker genes required to define the cell type cluster.


```{r}
library(randomForest)
set.seed(2)
dir.create("RF_figures")

target.type=as.character(tiss.subsamp@ident[1000])
target.type="Pancreas__pancreatic.A.cell"

markers <- read.csv('diffexpr_all.csv') # differential expression by Seurat, with subsampled cells
genes.use <- as.character(markers %>% 
                      filter(cluster==target.type,
                             gene %in% tf.names, 
                             p_val < 1e-3, 
                             avg_logFC > .2,
                             pct.1-pct.2 > 0.05,
                             pct.1 > 0.2) %>% 
                        arrange(desc(avg_logFC)) %>%
                      pull(gene))[1:100]
genes.use <- genes.use[!is.na(genes.use)]

train <- tiss.subsamp
response <- factor(tiss.subsamp@ident == target.type)
fit <- randomForest(x=t(as.matrix(tiss.subsamp@data[genes.use, ])),y=response,
                      importance=TRUE, 
                      ntree=100,mtry = length(genes.use))

# pdf(paste0("RF_figures/", target.type, "_varImpPlot.pdf"), height = 4, width = 6)
  varImpPlot(fit, n.var = 12)
# dev.off()
fit
plot(fit, log="y")

# Get top genes by fit
best.genes <- rownames(fit$importance)[order(fit$importance[,'MeanDecreaseAccuracy'], decreasing = T)][1:10]

require(rpart)
require(rpart.plot)
best.df <- data.frame(t(as.matrix(tiss.subsamp@data[best.genes[1:3], ])),y=response)
fit.n <- rpart(y ~ Arx + Fev, data = best.df, control = rpart.control(maxdepth = 3))
rpart.plot(fit.n)

# Can use this code to predict the class on another subsample of the dataset, or the whole dataset. 
# Prediction <- predict(fit, test)
# submit <- data.frame(PassengerId = test$PassengerId, Survived = Prediction)

require(party)
fit.1 <- ctree(y ~ Mafb + Arx + Fev, data = best.df,
           controls = ctree_control(stump = T))

fit.2 <- ctree(y ~ Mafb + Arx + Fev, data = best.df,
           controls = ctree_control(stump = T))

```

