---
title: "R Notebook"
output: html_notebook
---

```{r}
library(Seurat)
library(cowplot)
library(data.table)
library(dplyr)
library(tidyr)
library(parallel)
library(cba)
library(RColorBrewer)

tabula.dir <- "../../../tabula-muris/"
```

# Load Seurat object of data
```{r}
load("../../data/All_seurat_tiss.Robj")
```

# Load metadata and add to Seurat object
Loads Seurat object and list of TFs
Makes "annotation.2" metadata field which is tissue__cell_ontology_class. This is the field used to separate cells into classes for finding markers and doing correlations.
```{r}
metadata <- read.csv('../TM_facs_metadata.csv')
metadata <- metadata %>%  filter(cell_ontology_class!='unknown' & !is.na(tissue) & !is.na(cell_ontology_class))
metadata$tissue <- make.names(metadata$tissue)
metadata$cell_ontology_class <- make.names(metadata$cell_ontology_class)

# make "annotation.2"
metadata  <- metadata %>% mutate(annotation.2 = paste0(tissue, "__", cell_ontology_class))

tissue_colors <- read.csv(file.path(tabula.dir,'00_data_ingest/15_color_palette/tissue_colors.csv'))
colnames(tissue_colors) <- c('tissue','tiss.color')
tissue_colors$tissue <- make.names(tissue_colors$tissue)

metadata <- merge(metadata, tissue_colors, by = 'tissue')

rownames(metadata) <- metadata$cell # need this to add to Seurat object
metadata <- metadata[tiss@cell.names, ]

# Add metadata
tiss <- AddMetaData(tiss, metadata)
sum(is.na(tiss@meta.data$annotation.2))
length(unique(tiss@meta.data$annotation.2))
tiss <- SetAllIdent(tiss, 'annotation.2')

# Make gene names R compatible
rownames(tiss@data) <- make.names(rownames(tiss@data))
rownames(tiss@raw.data) <- make.names(rownames(tiss@raw.data))
rownames(tiss@scale.data) <- make.names(rownames(tiss@scale.data))
gc()
```


# Load TFs 
```{r}
# (1140 genes from MGI, filtered by GO Term = "DNA binding transcription factor activity", GO ID ?)
tfs      <- fread('../GO_term_summary_20171110_222852.csv')
tf.names <- as.character( tfs %>% distinct(Symbol) %>% pull(Symbol) )
tf.names <- make.names(tf.names)
tf.names <- tf.names[tf.names %in% rownames(tiss@data)]
length(tf.names)

# remove genes with very low expression in data
x=apply(as.matrix(tiss@data[tf.names, ]) > 0, 1, sum) 
genes.expr <- names(x[x > 5])
tf.names <- tf.names[tf.names %in% genes.expr]

# Remove IEGs
iegs <- as.character(read.csv(file.path(tabula.dir,"00_data_ingest/20_dissociation_genes/genes_affected_by_dissociation_unix.csv"), header = T)[,1])
tf.names <- tf.names[!tf.names %in% iegs]
length(tf.names)
```


# Make and save groups of cells 
For cross-tissue analysis of broad cell classes
```{r}
# endothelial
cell.type = "endothelial"
system(paste("mkdir",cell.type))
endo.labels <- tiss@meta.data %>% filter( (cell_ontology_class %like% 'endoth') | (cell_ontology_class %like% "endocard"))
endo.labels %>% distinct(tissue, cell_ontology_class, annotation.2)
endo.cells.all <- unique(as.character(endo.labels %>% pull(cell)))
write.csv(endo.cells.all, file = file.path(cell.type, 'cells_all.csv'), row.names = F)

# epithelial
cell.type = "epithelial"
system(paste("mkdir",cell.type))

epith.labels <- bind_rows(tiss@meta.data %>% filter(cell_ontology_class %like% 'epith'), 
                          tiss@meta.data %>% filter(cell_ontology_class %like% 'basal'),
                          tiss@meta.data %>% filter(cell_ontology_class %like% 'keratino'),
                           tiss@meta.data %>% filter(cell_ontology_class %like% 'epiderm'),
                          tiss@meta.data %>% filter(cell_ontology_class %like% 'urothelial'))
epith.labels %>% distinct(tissue, cell_ontology_class, annotation.2)
epith.cells.all <- unique(as.character(epith.labels %>% pull(cell)))
write.csv(epith.cells.all, file = file.path(cell.type, 'cells_all.csv'), row.names = F)



# stromal: leave as-is (separate by tissue and annotation)
cell.type = "stromal"
system(paste("mkdir",cell.type))

strom.labels <- tiss@meta.data %>% filter(cell_ontology_class %in% c('stromal', 'fibroblasts', 'mesenchymal','mesenchymal_stem','stromal_stem'))
strom.labels %>% distinct(tissue, cell_ontology_class)


strom.labels <- bind_rows(tiss@meta.data %>% filter(annotation.2 %like% 'strom'), 
                          tiss@meta.data %>% filter(annotation.2 %like% 'fibro'),
                          tiss@meta.data %>% filter(annotation.2 %like% 'mesen'))
strom.labels %>% distinct(tissue, cell_ontology_class, annotation.2)
strom.cells.all <- unique(as.character(strom.labels %>% pull(cell)))
write.csv(strom.cells.all, file = file.path(cell.type, 'cells_all.csv'), row.names = F)

```

# Functions
## Differential expression function
```{r}
# calculates positively-enriched TFs only
celltype.diffexpr <- function(data, cell.type, max.cells.per.ident=50, n.cores=2){
  cells.use <- as.character(read.csv(file.path(cell.type,'cells_all.csv'))[,1])
  
  print(head(cells.use))
  tiss.use <- SubsetData(data, cells.use = cells.use)
  tiss.use <- SetAllIdent(tiss.use, 'annotation.2')
  tiss.use <- SubsetData(tiss.use, max.cells.per.ident = max.cells.per.ident)

  print(table(tiss.use@ident))
  require(parallel)
  markers <- do.call(rbind, mclapply(unique(tiss.use@ident), function(x) {
    # In the version of Seurat for initial MACA pub (Seurat_2.0.1)
    # uses object@data as input for statistical test
    res=FindMarkers(tiss.use, ident.1 = x, only.pos = T, test.use="wilcox", min.diff.pct = 0.1, do.print=T, genes.use=tf.names)
    gc()
    res$cluster = x
    res$gene = rownames(res)
    return(res)}, mc.cores = n.cores))
  
  write.csv(tiss.use@cell.names,file.path(cell.type,'cells_subsamp.csv'), row.names=F)
  write.csv(markers,  file.path(cell.type,'diffexpr.csv'))
  
  rm(tiss.use, markers)
  gc()
}

```

## Correlogram function
```{r}

# returns output of barb.cormap (de)
plot.correlogram <- function(cell.type, pval_cutoff, avg_logFC_cutoff, nmarkers.per.celltype, height=10){
  markers <- read.csv(file.path(cell.type, 'diffexpr.csv'))
  markers$avg_logFC <- as.numeric(as.character(markers$avg_logFC))

  markers$p_val <- as.numeric(as.character(markers$p_val))

  genes.use <- unique(markers %>% group_by(cluster) %>% 
                      filter(gene %in% tf.names) %>%
                      filter(p_val < pval_cutoff, avg_logFC > avg_logFC_cutoff) %>% 
                      top_n(n=nmarkers.per.celltype, wt=avg_logFC) %>% 
                      pull(gene))

  
  cells.use = as.character(read.csv(file.path(cell.type, 'cells_subsamp.csv'))[,1])
  tiss.use <- SubsetData(tiss, cells.use = cells.use)
  tiss.use <- SetAllIdent(tiss.use, 'annotation.2')

  # plot heatmap of top cell-type-specific genes
  print(paste("number of genes: ", length(genes.use)))
  dt.plot <- t(as.matrix(tiss.use@data[genes.use, ]))

  rownames(dt.plot) <- tiss.use@ident
  # res <- cellorder.heatmap(dt.plot, file.path(cell.type, 'TF_cellHeatmap.pdf'), height = 15, width = 18, cexCol = .8, cexRow = .1)
  
  # Calculate correlations
  mat.use <- t(as.matrix(tiss.use@data[genes.use, ]))
  tf.cor    <- cor(mat.use)
  
  enrich.score <- dcast(markers, gene ~ cluster, value.var = 'avg_logFC')
  rownames(enrich.score) <- enrich.score[,"gene"]
  enrich.score <- enrich.score[, 2:ncol(enrich.score)]
  # change rownames to the cell type and tissue that each gene is enriched in (ordering by avg_logFC above)
  topIDenriched   <- sapply(colnames(tf.cor), function(x) {
    names(sort(t(enrich.score)[, x],  decreasing = T))[1]}) 
  colnames(tf.cor) <- topIDenriched
  
  correlo.out <- barb.cormap(tf.cor,  file.path(cell.type, 'TF_cormap.pdf'), height =height, width = height)
  
  # Generate row colors of heatmap corresponding to highest-expressing cell type
  topID.plotorder <- correlo.out[[2]]
  meta.summary <- tiss.use@meta.data %>% distinct(tissue, cell_ontology_class, tiss.color, annotation.2)
  # print(head(meta.summary))
  ntypes=length(unique(meta.summary$cell_ontology_class))
  tmp <- colorRampPalette(brewer.pal(min(ntypes, 11), 'Paired'))(ntypes)
  annot_colors <- data.frame(annot.colors = tmp, cell_ontology_class = unique(meta.summary$cell_ontology_class))
  write.csv(annot_colors, file=file.path(cell.type, 'annotColors.csv'))
  
  # print(head(annot_colors))
  meta.summary <- merge(meta.summary, annot_colors, by = 'cell_ontology_class')
  # print(head(meta.summary))

  plot.colors <- data.frame(annotation.2 = topID.plotorder)
  # print(head(plot.colors))

  plot.colors <- merge(plot.colors, meta.summary %>% select(annotation.2, annot.colors, tiss.color), by = 'annotation.2')
  # print((plot.colors))
  plot.colors <- plot.colors[match(topID.plotorder, plot.colors$annotation.2), ]
  # print((plot.colors))
  plot.colors$ymin <- 0.1*(0:(nrow(plot.colors)-1))
  plot.colors$ymax <- 0.1*(1:(nrow(plot.colors)))
  # print((plot.colors))
  plot.colors$tiss.color <- toupper(plot.colors$tiss.color)
    plot.colors$annot.colors <- toupper(plot.colors$annot.colors)


  require(grDevices)
  pdf( file.path(cell.type, 'TF_cormap_rowcolors.pdf'), height = 10, width = 4)
    plot(c(0, 2), c(0, max(plot.colors$ymax) + 1), type = "n", xlab = "", ylab = "",
         main = "plot colors")
    rect(0,plot.colors$ymin, 1 , plot.colors$ymax, col = plot.colors$annot.colors, border = NA)
    rect(1,plot.colors$ymin, 2 , plot.colors$ymax, col = plot.colors$tiss.color, border = NA)
  dev.off()

  return(correlo.out)
}

# plotting function for correlogram
barb.cormap <- function(mat.cor,fname,width=12,height=12,method="complete",cex=0.5,mincor=-1,maxcor=1){
  require(lattice)
  require(cba)
  rowdist <- dist(mat.cor)
  coldist <- dist(mat.cor, by_rows = F)
  hc.cor <- hclust(coldist, method=method)
  hr.cor <- hclust(rowdist, method=method)

  optimal.row <- order.optimal(rowdist,hr.cor$merge)
  optimal.col <- order.optimal(coldist,hc.cor$merge)
 
  ord.row <- optimal.row$order
  ord.col <- optimal.col$order
  
  plt = levelplot(mat.cor[ord.row,ord.col],xlab=NULL,ylab=NULL,
                  at=do.breaks(c(mincor-0.01,maxcor+0.01),19),scales=list(x=list(rot=90),cex=cex),
                  colorkey=list(space="top"),
                  col.regions=colorRampPalette(c("dodgerblue4", "dodgerblue", "white", "lightcoral", "firebrick4"), space="Lab"))
  pdf(fname,width=width,height=height)
    print(plt)
  dev.off()
  
  return(list(rownames(mat.cor[ord.row, ]), colnames(mat.cor[, ord.col]), plt, hc.cor, hr.cor))
}

```

## tSNE plotting functions
```{r}
make.tSNE <- function(data, cell.type, n.pcs=10, calc.PCA.only = F, subset.cells = F){
  
  dir=cell.type
  if(!file.exists(dir)) system(paste0('mkdir ',dir))

  cells.use <- as.character(read.csv(file.path(cell.type,'cells_all.csv'))[,1])
  tiss.use <- SubsetData(data, cells.use = cells.use)
  tiss.use <- SetAllIdent(tiss.use, 'annotation.2')

  if(subset.cells){
    cells.use = as.character(read.csv(paste0('cells_',cell.type,'.csv'))[,1])
    tiss.use <- SubsetData(tiss, cells.use = cells.use)
  }
  
  tiss.use <- FindVariableGenes(object = tiss.use, do.plot = TRUE, x.high.cutoff = Inf, y.cutoff = 0.5)
  tiss.use@var.genes <- tiss.use@var.genes[!tiss@var.genes %in% iegs]

  tiss.use <- RunPCA(object = tiss.use, do.print = FALSE)
  tiss.use <- ProjectPCA(object = tiss.use, do.print = FALSE)
  print(PCElbowPlot(object = tiss.use))

  if(!calc.PCA.only){
    tiss.use <- RunTSNE(object = tiss.use, dims.use = 1:n.pcs, seed.use = 10, perplexity=30)
    TSNEPlot(object = tiss.use, do.label = T)
  }
  
  write.csv(GetCellEmbeddings(tiss.use, reduction.type = 'tsne'), file = file.path(cell.type,'tsne.csv'))

}

plot.tsne.annot <- function(cell.type){
  dir=cell.type
  tsne <- fread(file.path(dir, 'tsne.csv'))
  setnames(tsne,  c('cell.name','tSNE_1','tSNE_2'))

  meta.plot <- merge(tsne, tiss@meta.data, by = 'cell.name')

  p=ggplot(meta.plot, aes(tSNE_1, tSNE_2, color = tiss.color)) + 
    geom_point() +
    coord_fixed() +
    scale_color_identity()+
    theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank())
  save_plot(file.path(cell.type,'tsne_tissue.pdf'),p)

  p=ggplot(meta.plot, aes(tSNE_1, tSNE_2, color = annotation)) + 
    geom_point() +
    coord_fixed() +
    scale_color_brewer(palette = 'Paired') +
    theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank())
    save_plot(file.path(cell.type,'tsne_annotation.pdf'),p)

}

plot.tsne.markers <- function(cell.type, pval_cutoff, avg_logFC_cutoff, nmarkers.per.celltype){

  # plot tsne with tissue colors
  tsne <- fread(file.path(cell.type,'tsne.csv'))
  setnames(tsne,  c('cell.name','tSNE_1','tSNE_2'))

  metadata <- as.data.table(tiss@meta.data)

  markers <- read.csv(paste0('diffexpr_',cell.type,'.csv'))
  markers <- markers %>% filter(!gene %like% "-")
  genes.use <- as.character(unique(markers %>% group_by(cluster) %>% 
                        filter(gene %in% tf.names) %>%
                        filter(p_val < pval_cutoff, avg_logFC > avg_logFC_cutoff) %>% 
                        top_n(n=nmarkers.per.celltype, wt=avg_logFC) %>% 
                        pull(gene)))
  
  expr.data <- data.table(cell.name=tsne$cell.name, t(as.matrix(tiss@data[genes.use, tsne$cell.name])))
  tsne.genes <- merge(tsne, expr.data, by='cell.name')

  genes.fixednames <- gsub("-", ".", genes.use)
  genes.translate <- data.frame(gene.fixed=genes.fixednames, row.names = genes.use)
  genes.translate$gene.fixed <- as.character(genes.translate$gene.fixed)
  plt.list <- lapply(genes.use, function(x) {
    p=ggplot(tsne.genes, aes_string('tSNE_1', 'tSNE_2', color=genes.translate[x, 'gene.fixed'])) + 
    geom_point() +
    scale_color_gradient(low = 'grey80', high='firebrick4') +
    theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank(),
          plot.title = element_text(size = 35, face = "bold")) +
      ggtitle(x)
    return(p)
  })
  
  plt=plot_grid(plotlist = plt.list, align = 'hv', nrow=ceiling(sqrt(length(genes.use))), ncol = ceiling(sqrt(length(genes.use))))
  save_plot(file.path(cell.type,'tsne_markerGenes.pdf'), plt,  nrow=ceiling(sqrt(length(genes.use))), ncol = ceiling(sqrt(length(genes.use))))

}

```


# Calculate differentially-expressed genes across tissues w/in a broad cell class.
Cells are subsampled here which speeds up the calculation. The subsampled cells are saved and used to generate the correlation plots in the next function. This next part is critical because the correlation is done using individual cells as datapoints and not across cell type averages, so the cells must be subsampled to avoid subtypes with many cells overtaking correlation. 
```{r}
celltype.diffexpr(data = tiss, cell.type = 'endothelial', max.cells.per.ident = 60, n.cores = 4)
celltype.diffexpr(data = tiss, cell.type = 'epithelial', max.cells.per.ident = 60, n.cores = 4)
celltype.diffexpr(data = tiss, cell.type = 'stromal', max.cells.per.ident = 100, n.cores = 4)
```


```{r}
# low pval/expr diff cutoffs
corr.endo=plot.correlogram('endothelial', 1e-7, 0.3, 8, height=8)

corr.epith=plot.correlogram('epithelial', 1e-7, 0.35, 8)

corr.strom=plot.correlogram('stromal', 1e-7, 0.3, 8)
 
```

