---
title: "Brain_Non-Myeloid facs Supplementary Notebook"
output: html_notebook
---

```{r}
library(here)
source(here('30_tissue_supplement_figures', 'supplemental_figures.R'))
save_folder = here('30_tissue_supplement_figures', 'Brain_Non-Myeloid', 'facs')
dir.create(save_folder, recursive=TRUE)

tissue_of_interest = 'Brain_Non-Myeloid'
filename = paste0('facs_',tissue_of_interest, '_seurat_tiss.Robj')
load(here('00_data_ingest', '04_tissue_robj_generated', filename))

additional.group.bys = sort(c("subtissue"))

group.bys = c(standard.group.bys, additional.group.bys)

genes_to_check = c("A2m", "Aldh1l1", "Aqp4", "Gdf10", "Naga", "Nbl1", "Nsdhl", "Slc1a3", "St6galnac5", "Tnfaip2")
dot_tsne_violin(tiss, genes_to_check, save_folder, prefix = prefix, group.bys)
```


```{r}
tiss.markers <- FindAllMarkers(object = tiss, only.pos = TRUE, min.pct = 0.25, thresh.use = 0.25)
filename = file.path(save_folder, paste(prefix, 'findallmarkers.csv', sep='_'))
```



## To identify excititory, inhibitory, and NPCs
```
genes_to_check = c('Slc17a7','Neurod6','Mab21l1','Gad1','Reln','Calb1','Dcx','Dlx2','Ascl1','Hes5')
filename = file.path(save_folder, paste(prefix, group.by, 'violinplot.pdf', sep='_'))
  VlnPlot(tiss, genes_to_check, group.by = column, do.return=TRUE)
  ggsave(filename, dpi=300)
  dev.off()
```

# Subsetting of brain endothelial cells and analysis using robust PCA followed by Seurat

Filter for all endothelial cells from whole brain (non-microglia) annotation
```{r}
endo.sub<-tiss@raw.data[, grep("endothelial",tiss@meta.data$cell_ontology_class)]
endo<-grep("endothelial",tiss@meta.data$cell_ontology_class)
endo.sub.names<-colnames(tiss@scale.data)[endo]
endo.sub<-tiss@raw.data[,endo.sub.names]

```

Load packages
```{r}
library(dplyr)
library(Matrix)
library(cowplot)
require(RColorBrewer)
```

Required custom functions
```{r}

# return minimum (most negative) correlation value for each gene
get.gene.cor <- function(data, min.cells.expr = 1) {
  mat = t(data[, colSums(data > 0) > min.cells.expr]) # filter genes
  mat = mat - rowMeans(mat)
  cr = tcrossprod( mat/sqrt(rowSums(mat^2)) )
  return(cr)
}
get.min.cor <- function(data, min.cells.expr = 1) {
  require(matrixStats)
  cr = get.gene.cor(data, min.cells.expr)
  min.cor = colMins(cr, na.rm =T)
  names(min.cor) = colnames(cr)
  return(min.cor)
}


## robust PCA function
do.robpca <- function(data.counts,min_cells=0,min_sd=0,ncp=10,expression.values="log10.cpm",from.mat = T,...){
  require(rrcov)
  if(!from.mat){
    if(min_cells>0){
      data<-filter.genes(data.counts,expression.values,min_sd,min_cells)
    } else{ cast.counts <- cast.to.df(data.counts,expression.values) }
  } else {
    cast.counts = data.counts[, colSums(data.counts > 0) > min_cells]
  }
  print('starting rPCA')
  pca <- PcaHubert(cast.counts, k = ncp,kmax=ncp,...)
  rm(cast.counts);gc()
  scores <- as.data.frame(getScores(pca));scores$cell.name <- rownames(scores); scores <- data.table(scores);setkey(scores)
  var   <- as.data.frame(getLoadings(pca)); var$gene <- rownames(var);var <- as.data.table(var);setkey(var)
  gc()
  list(pca, scores, var)
}

filter.genes <- function(data.counts,expression.value="log10.cpm",min_sd=0,min_cells=0){
  data=copy(data.counts);
  data[,numcells_exp:=sum(get(expression.value)>0),by=gene]; data<-data[numcells_exp>min_cells]
  data[,gene_sd:=sd(get(expression.value)),by=gene]; data<-data[gene_sd>min_sd]
  data
}


## cast to dataframe function
cast.to.df <- function(data.counts,expression.values="log10.cpm",annot="none", to.matrix = F, genes.use = NULL){
  if(!is.null(genes.use)) data.counts = data.counts[gene %in% genes.use]
  if(annot != "none"){
    cast.counts <- dcast.data.table(data.counts, cell.name ~ gene, value.var=expression.values, fill = 0)
    setkey(data.counts)
    cast.counts <- merge(cast.counts,unique(data.counts[,.(cell.name,get(annot))]),by="cell.name")
    setnames(cast.counts,c(colnames(cast.counts)[1:(ncol(cast.counts)-1)],annot))
    cast.counts <- as.data.frame(cast.counts)
  }
  else{
    cast.counts <- as.data.frame(dcast.data.table(data.counts, cell.name ~ gene, value.var=expression.values,fill=0))
    gc()
  }
  rownames(cast.counts)<-cast.counts$cell.name; cast.counts <- cast.counts[,c(ncol(cast.counts),2:(ncol(cast.counts)-1))]
  if(to.matrix) cast.counts = as.matrix(cast.counts)
  cast.counts
}


```



QC
```{r}
ECs<-CreateSeuratObject(raw.data=endo.sub, min.cells=5, min.genes=5) # already filter zero count genes
colnames(ECs@meta.data)[colnames(ECs@meta.data)=='nUMI']<-'nReads' ##change name

#Add ribo data
ribo.genes<-grep(pattern="^Rp[s1][[:digit:]]", x=rownames(x=ECs@data), value=TRUE) ##60
percent.ribo<-Matrix::colSums(ECs@raw.data[ribo.genes, ])/Matrix::colSums(ECs@raw.data)
ECs <- AddMetaData(object = ECs, metadata = percent.ribo, col.name = "percent.ribo")

#Add mito data
mito.genes <- grep(pattern = "^mt-", x = rownames(x = ECs@data), value = TRUE)
percent.mito <- Matrix::colSums(ECs@raw.data[mito.genes, ])/Matrix::colSums(ECs@raw.data)
ECs <- AddMetaData(object = ECs, metadata = percent.mito, col.name = "percent.mito")

GenePlot(object = ECs, gene1 = "nReads", gene2 = "nGene")

ECs <- FilterCells(object = ECs, subset.names = c("nGene", "nReads"), low.thresholds = c(200, 20000), high.thresholds = c(25000, Inf))
ECs <- NormalizeData(object = ECs)
ECs <- ScaleData(object = ECs)

```


Add brain region information in metadata
```{r}

DF<-as.data.frame(cbind(tiss@cell.names, as.character(tiss@meta.data$subtissue)))
rownames(DF)<-DF$V1
ECs@meta.data$subtissue<-as.character(DF[ECs@cell.names, 2])

```


Feature selection: perform robust PCA on top 2500 most anti-correlated genes

```{r}

data.pca <- t(log2(1+1e6*(t(t(ECs@raw.data)/colSums(ECs@raw.data))))) ##log 2 normalize data
gene.min.cors <- get.min.cor(data.pca, min.cells.expr = 2)
genes.use <- names(sort(gene.min.cors)[1:2500])

# Remove some housekeeping genes
genes.use <- genes.use[grep('^Rp[ls].*', genes.use, invert = T)]
genes.use <- genes.use[grep('Rn45s', genes.use, invert = T)]
genes.use <- genes.use[grep('Lars2', genes.use, invert = T)]
genes.use <- genes.use[grep('Malat1', genes.use, invert = T)]

ncp=20
num.sig.genes <- 30

pcalist <- do.robpca(data.pca[, genes.use],ncp = ncp, from.mat = T)
loadings <- pcalist[[3]]
sdev<-pcalist[[1]]
sdev<-sdev@sd
data.scaled<-t(t(data.pca[,genes.use]) / colMaxs(data.pca[,genes.use]))

library(pbapply)
dims <- paste0('PC', 1:ncp) ## 1:20
pc.sigs <- as.data.frame(do.call(cbind, lapply(dims, function(dim) {
  setorderv(loadings,dim,-1)
  genes.pos <- loadings[,gene][1:num.sig.genes]
  setorderv(loadings,dim,1)
  genes.neg <- loadings[,gene][num.sig.genes:1]
  cast.sig <- as.data.frame(cbind(rowSums(data.scaled[, genes.pos]), rowSums(data.scaled[, genes.neg])))
  print(head(cast.sig))
  cast.sig$pc.score <- cast.sig[, 1] - cast.sig[, 2]
  colnames(cast.sig) <- c(paste0(dim,'.pos'), paste0(dim,'.neg'), dim)
  return(as.matrix(cast.sig))
  }
)))

pc.sigs <- pc.sigs[match(ECs@cell.names, rownames(pc.sigs)), ]
pc.sigs <- as.matrix(pc.sigs[, grepl("PC[0-9]+$", colnames(pc.sigs))])
ECs <- SetDimReduction(ECs, 'pca', 'cell.embeddings', new.data = pc.sigs)
pc.loadings<-loadings[,1:20]
pc.loadings<-as.matrix(pc.loadings)
rownames(pc.loadings)<-loadings$gene
ECs <- SetDimReduction(ECs, 'pca', 'gene.loadings', new.data = pc.loadings)
ECs@dr$pca@key<-"PC"

```



Plug the cell.embeddings and gene.loadings from rPCA back into Seurat object and run Seurat clustering
```{r}

ECs <- FindClusters(object = ECs, reduction.type = "pca", dims.use = 1:15,
    resolution = 1, print.output = 0, save.SNN = TRUE, force.recalc = T)

ECs <- RunTSNE(object = ECs, dims.use = 1:8, seed.use = 10, check_duplicates = F, perplexity=40)
  
# unsupervised clustering
filename = file.path(save_folder, paste('endothelial', 'cluster_number', 'tsneplot.png', sep='_'))
TSNEPlot(object = ECs, do.label = T, label.size = 6, pt.size = 3.1, colors.use = brewer.pal(11,'Paired'), do.return=TRUE) + labs(x = "tSNE 1", y="tSNE 2")+ theme(axis.title.x = element_text(colour = "black", size=16), axis.title.y = element_text(colour = "black", size=16), legend.text=element_text(size=13))
ggsave(filename, dpi=300)
dev.off()

#annotated by brain subregion
filename = file.path(save_folder, paste('endothelial', 'subtissue, 'tsneplot.png', sep='_'))
TSNEPlot(object = ECs, do.label = T, label.size = 6, pt.size = 3.1, colors.use = brewer.pal(9,'Set1'), do.return=TRUE, group.by="subtissue") + labs(x = "tSNE 1", y="tSNE 2")+ theme(axis.title.x = element_text(colour = "black", size=16), axis.title.y = element_text(colour = "black", size=16), legend.text=element_text(size=13))
ggsave(filename, dpi=300)
dev.off()

PCHeatmap(object = ECs, pc.use = 1:12, cells.use = 300, do.balanced = TRUE, label.columns = FALSE, num.genes = 20)
PCAPlot(object = ECs, dim.1 = 1, dim.2 = 2, pt.size = 2)


#violin plots of key defining genes for the Inflamed (Venous) and Notch (Arterial) populations
genes_check<-c('Vcam1','Icam1','Lcn2','Hif1a','Vwf','Csf1','Notch1','Hey1','Vegfc','Edn1','Tmem100')

filename = file.path(save_folder, paste('endothelial', 'subtissue, 'featureplot.png', sep='_'))
FeaturePlot(ECs, features.plot=genes_check, pt.size = 2, no.axes = T, cols.use = c("lightgray", "red"), dark.theme = F)
ggsave(filename, dpi=300)
dev.off()

filename = file.path(save_folder, paste('endothelial', 'subtissue, 'violinplot.png', sep='_'))
VlnPlot(object = ECs, features.plot = genes_check, do.sort=F, size.title.use = 18, size.x.use = 0.3,size.y.use = 3, cols.use = brewer.pal(11,'Paired'), point.size.use = 0.3)
ggsave(filename, dpi=300)
dev.off()

```


Find differentially expressed genes between clusters
```{r}


DE_markers <- mclapply(levels(ECs@ident),  function(x) {
  df <- FindMarkers(object = ECs, ident.1 = x, only.pos = TRUE, min.pct = 0.20, thresh.use = 0.20)
  df$gene <- rownames(df)
  df$cluster <- x
  return(df)
  }, mc.cores = 2)
DE_markers <- bind_rows(DE_markers)

DE_markers <- DE_markers[DE_markers$p_val<0.05, ]
DE_markers_sorted <- DE_markers %>% group_by(cluster) %>% top_n(10, avg_diff)


DoHeatmap(object = ECs, genes.use = DE_markers_sorted$gene, slim.col.label = TRUE, remove.key = FALSE, cex.row = 7)


```



#renaming cluster for function AND segmental identity based on differentially expressed genes
```{r}

current.cluster.ids <- c(0, 1, 2, 3, 4)
new.cluster.ids <- c("Capillary/BBB-maintenance", "Capillary/BBB-maintenance","Arterial/Notch-signaling", "Capillary/BBB-maintenance", "Venous/inflammatory")
ECs@ident <- plyr::mapvalues(x = ECs@ident, from = current.cluster.ids, to = new.cluster.ids)
ECs<-StashIdent(object=ECs, save.name="Annotation") ## for more clusters

filename = file.path(save_folder, paste('endothelial', 'function_and_vessel_type', 'tsneplot.png', sep='_'))
TSNEPlot(object = ECs, do.label = T, pt.size = 3, colors.use = brewer.pal(8,'Set2'), do.return=TRUE) + labs(x = "tSNE 1", y="tSNE 2")+ theme(axis.title.x = element_text(colour = "black", size=16), axis.title.y = element_text(colour = "black", size=16), legend.text=element_text(size=13))
ggsave(filename, dpi=300)
dev.off()

```


Make Dot plot based top defining genes for each annotated endothelial cluster
```{r}
filename = file.path(save_folder, paste('endothelial', 'function_and_vessel_type', 'featureplot.png', sep='_'))
features.plot <- c("Vwf","Flt4","Nr2f2","Ephb4","Car4","Slc16a1","Tfrc","Efnb2","Jag1","Bmx")
ggsave(filename, dpi=300)
dev.off()

filename = file.path(save_folder, paste('endothelial', 'function_and_vessel_type', 'dotplot.png', sep='_'))
DotPlot(object = ECs, genes.plot = features.plot, plot.legend = TRUE,group.by = "Annotation", x.lab.rot = T)
ggsave(filename, dpi=300)
dev.off()

```

## Subset: ASTROCYTES

```{r}
ASTROCYTES.cells.use = rownames(tiss@meta.data)[grep("astrocyte",tiss@meta.data$cell_ontology_class)]
ASTROCYTES.n.pcs = 10
ASTROCYTES.res = 1
ASTROCYTES.genes_to_check = c("A2m", "Aldh1l1", "Aqp4", "Gdf10", "Naga", "Nbl1", "Nsdhl", "Slc1a3", "St6galnac5", "Tnfaip2")
ASTROCYTES.tiss = SubsetData(tiss, cells.use=ASTROCYTES.cells.use, )
ASTROCYTES.tiss <- ASTROCYTES.tiss %>% ScaleData() %>% 
  FindVariableGenes(do.plot = TRUE, x.high.cutoff = Inf, y.cutoff = 0.5) %>%
  RunPCA(do.print = FALSE)
ASTROCYTES.tiss <- ASTROCYTES.tiss %>% FindClusters(reduction.type = "pca", dims.use = 1:sub.n.pcs, 
    resolution = sub.res.use, print.output = 0, save.SNN = TRUE) %>%
    RunTSNE(dims.use = 1:sub.n.pcs, seed.use = 10, perplexity=30)

# Append this subset's groupby to the list
group.bys = c(group.bys, ""subtissue"")dot_tsne_violin({name}.tiss, {name}.genes_to_check,
    save_folder, prefix = {name}, group.bys)

```