---
title: "R Notebook"
output: html_notebook
---

```{r}
library(Seurat)
library(cowplot)
library(data.table)
library(dplyr)
library(tidyr)
library(parallel)
library(cba)
library(here)
library(RColorBrewer)
library(tidyverse)
library(openxlsx)
library(gplots)
library(reshape2)

library(ggplot2)
library(ggdendro)
library(scales)
library(wordspace)
library(GMD)

# library(devtools)
# install_github("cvarrichio/Matrix.utils")
library(Matrix.utils)
```

# Load Seurat object of data
```{r}
#tiss <- load(here('00_data_ingest', '11_global_robj','FACS_all.Robj'))
tm.droplet.matrix = readRDS(here("data-tabula-muris", "TM_droplet_mat.rds"))
tm.droplet.metadata = read_csv(here("data-tabula-muris", "TM_droplet_metadata.csv"))
tissDROPLET <- CreateSeuratObject(raw.data = tm.droplet.matrix, meta.data = tm.droplet.metadata)

tm.facs.matrix = readRDS(here("data-tabula-muris", "TM_facs_mat.rds"))
tm.facs.metadata = read_csv(here("data-tabula-muris", "TM_facs_metadata.csv"))
tissFACS <- CreateSeuratObject(raw.data = tm.facs.matrix, meta.data = tm.facs.metadata)
```

<!-- # Load metadata and add to Seurat object -->
<!-- Loads Seurat object and list of TFs -->
<!-- Makes "annotation.2" metadata field which is tissue__cell_ontology_class. This is the field used to separate cells into classes for finding markers and doing correlations. -->
<!-- ```{r} -->
<!-- metadata <- read.csv('../TM_facs_metadata.csv') -->
<!-- metadata <- metadata %>%  filter(cell_ontology_class!='unknown' & !is.na(tissue) & !is.na(cell_ontology_class)) -->
<!-- metadata$tissue <- make.names(metadata$tissue) -->
<!-- metadata$cell_ontology_class <- make.names(metadata$cell_ontology_class) -->

<!-- # make "annotation.2" -->
<!-- metadata  <- metadata %>% mutate(annotation.2 = paste0(tissue, "__", cell_ontology_class)) -->

<!-- # tissue_colors <- read.csv(file.path(tabula.dir,'00_data_ingest/15_color_palette/tissue_colors.csv')) -->
<!-- tissue_colors <- read.csv(here('00_data_ingest', '15_color_palette','tissue_colors.csv')) -->
<!-- colnames(tissue_colors) <- c('tissue','tiss.color') -->
<!-- tissue_colors$tissue <- make.names(tissue_colors$tissue) -->

<!-- metadata <- merge(metadata, tissue_colors, by = 'tissue') -->

<!-- rownames(metadata) <- metadata$cell # need this to add to Seurat object -->
<!-- metadata <- metadata[tiss@cell.names, ] -->

<!-- # Add metadata -->
<!-- tiss <- AddMetaData(tiss, metadata) -->
<!-- sum(is.na(tiss@meta.data$annotation.2)) -->
<!-- length(unique(tiss@meta.data$annotation.2)) -->
<!-- tiss <- SetAllIdent(tiss, 'annotation.2') -->

<!-- # Make gene names R compatible -->
<!-- rownames(tiss@data) <- make.names(rownames(tiss@data)) -->
<!-- rownames(tiss@raw.data) <- make.names(rownames(tiss@raw.data)) -->
<!-- rownames(tiss@scale.data) <- make.names(rownames(tiss@scale.data)) -->
<!-- gc() -->
<!-- ``` -->

# Load TFs 
```{r}
# (1140 genes from MGI, filtered by GO Term = "DNA binding transcription factor activity", GO ID ?)
tfs      <- read.csv(here('23_tf_analysis','GO_term_summary_20171110_222852.csv'))
tf.names <- as.character( tfs %>% distinct(Symbol) %>% pull(Symbol) )
tf.names <- make.names(tf.names)
tf.names <- tf.names[tf.names %in% rownames(tissFACS@data)]
length(tf.names)

# remove genes with very low expression in data
x=apply(as.matrix(tissFACS@data[tf.names, ]) > 0, 1, sum) 
genes.expr <- names(x[x > 5])
tf.names <- tf.names[tf.names %in% genes.expr]

# Remove IEGs
# iegs <- as.character(read.csv(file.path(tabula.dir,"00_data_ingest/20_dissociation_genes/genes_affected_by_dissociation_unix.csv"), header = T)[,1])

iegs <- as.character(read.csv(here('00_data_ingest','20_dissociation_genes','genes_affected_by_dissociation_unix.csv'), header = T)[,1])

tf.names <- tf.names[!tf.names %in% iegs]
length(tf.names)
```

# Calculate global dendogram
```{r}
# tm.facs.matrix.tfs <- tm.facs.matrix[rownames(tm.facs.matrix) %in% unique(tfs$Symbol),]

tm.facs.matrix.tfs <- tm.facs.matrix[rownames(tm.facs.matrix) %in% tf.names,]
tm.facs.matrix.tfs <- as.data.frame(t(as.matrix(tm.facs.matrix.tfs)))


tm.facs.matrix.tfs$factors <- as.factor(tm.facs.metadata$cell_ontology_class)
#tm.facs.matrix.tfs$tissues <- as.factor(tm.facs.metadata$tissue)

# consider only TFs that are not affected by dissociation
tm.facs.matrix.tfs.cellids <- tm.facs.matrix.tfs %>% 
  group_by(factors) %>%
  summarise_at(vars(colnames(tm.facs.matrix.tfs[1:1016])),funs(sum(., na.rm=TRUE)))


tm.facs.matrix.tfs.cellids.norm = tm.facs.matrix.tfs.cellids
for (i in 1:82){
  jaux = sum(tm.facs.matrix.tfs.cellids[i,2:1017])
  for (j in 2:1017){
    tm.facs.matrix.tfs.cellids.norm[i,j] = tm.facs.matrix.tfs.cellids[i,j]/jaux
  }
}
tm.facs.matrix.tfs.cellids.norm <- tm.facs.matrix.tfs.cellids.norm[,-1]

hmcol = colorRampPalette(brewer.pal(9, "RdBu"))(100)
heatmap.2(as.matrix(as.data.frame(lapply(tm.facs.matrix.tfs.cellids.norm, as.numeric))),scale="column",trace = "none",col = hmcol, dendrogram = "row", labRow = tm.facs.matrix.tfs.cellids$factors, labCol = FALSE)

d <- dist(as.matrix(as.data.frame(lapply(tm.facs.matrix.tfs.cellids[,-1], as.numeric))), method = "euclidean", diag = FALSE, upper = FALSE, p = 2)
h <- hclust(d, method = "complete", members = NULL)
plot(h, labels = tm.facs.matrix.tfs.cellids$factors, hang = 0.1, check = TRUE,
     axes = TRUE, frame.plot = FALSE, ann = FALSE,
     main = "Cluster Dendrogram",
     sub = NULL, xlab = "", ylab = "", which.plots=2, cex=.5)


dnorm <- dist(as.matrix(as.data.frame(lapply(tm.facs.matrix.tfs.cellids.norm, as.numeric))), method = "euclidean", diag = FALSE, upper = FALSE, p = 2)
hnorm <- hclust(dnorm, method = "complete", members = NULL)
plot(hnorm, labels = tm.facs.matrix.tfs.cellids$factors, hang = 0.1, check = TRUE,
     axes = TRUE, frame.plot = FALSE, ann = FALSE,
     main = "Cluster Dendrogram",
     sub = NULL, xlab = "", ylab = "", which.plots=2, cex=.5)

#convert cluster object to use with ggplot
dendr <- dendro_data(h, type="rectangle") 

#your own labels (now rownames) are supplied in geom_text() and label=label
ggplot() + 
  geom_segment(data=segment(dendr), aes(x=x, y=y, xend=xend, yend=yend)) + 
  geom_text(data=label(dendr), aes(x=x, y=y, label=tm.facs.matrix.tfs.cellids$factors, hjust=0), size=3) +
  coord_flip() + scale_y_reverse(expand=c(0.2, 0)) + 
  theme(axis.line.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.y=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_rect(fill="white"),
        panel.grid=element_blank())
ggsave("tf_dendro_all.pdf", plot = last_plot(), device = "pdf", path = NULL,
  scale = 1, width = 20, height = 20, units = "cm",
  dpi = 150, limitsize = FALSE)

```

```{r}

# plot the global expression of TFs as histogram (let's try!)
hist(as.matrix(as.data.frame(lapply(tm.facs.matrix.tfs.cellids, as.numeric))))
hist(as.matrix(as.data.frame(lapply(tm.facs.matrix, as.numeric))))


```


```{r}
# cnsider all TFs
tm.facs.matrix.tfs.cellids <- tm.facs.matrix.tfs %>% 
  group_by(factors) %>%
  summarise_at(vars(colnames(tm.facs.matrix.tfs[1:1048])),funs(sum(., na.rm=TRUE)))

tm.facs.matrix.tfs.cellids.norm = tm.facs.matrix.tfs.cellids
for (i in 1:82){
  jaux = sum(tm.facs.matrix.tfs.cellids[i,2:1049])
  for (j in 2:1049){
    tm.facs.matrix.tfs.cellids.norm[i,j] = tm.facs.matrix.tfs.cellids[i,j]/jaux
  }
}
tm.facs.matrix.tfs.cellids.norm <- tm.facs.matrix.tfs.cellids.norm[,-1]

heatmap.2(as.matrix(as.data.frame(lapply(tm.facs.matrix.tfs.cellids.norm, as.numeric))),scale="row",trace = "n",)

d <- dist(as.matrix(as.data.frame(lapply(tm.facs.matrix.tfs.cellids[,-1], as.numeric))), method = "euclidean", diag = FALSE, upper = FALSE, p = 2)
h <- hclust(d, method = "complete", members = NULL)
plot(h, labels = tm.facs.matrix.tfs.cellids$factors, hang = 0.1, check = TRUE,
     axes = TRUE, frame.plot = FALSE, ann = FALSE,
     main = "Cluster Dendrogram",
     sub = NULL, xlab = "", ylab = "", which.plots=2, cex=.5)


dnorm <- dist(as.matrix(as.data.frame(lapply(tm.facs.matrix.tfs.cellids.norm, as.numeric))), method = "euclidean", diag = FALSE, upper = FALSE, p = 2)
hnorm <- hclust(dnorm, method = "complete", members = NULL)
plot(hnorm, labels = tm.facs.matrix.tfs.cellids$factors, hang = 0.1, check = TRUE,
     axes = TRUE, frame.plot = FALSE, ann = FALSE,
     main = "Cluster Dendrogram",
     sub = NULL, xlab = "", ylab = "", which.plots=2, cex=.5)


#convert cluster object to use with ggplot
dendr <- dendro_data(h, type="rectangle") 

#your own labels (now rownames) are supplied in geom_text() and label=label
ggplot() + 
  geom_segment(data=segment(dendr), aes(x=x, y=y, xend=xend, yend=yend)) + 
  geom_text(data=label(dendr), aes(x=x, y=y, label=tm.facs.matrix.tfs.cellids$factors, hjust=0), size=3) +
  coord_flip() + scale_y_reverse(expand=c(0.2, 0)) + 
  theme(axis.line.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.y=element_blank(),
        axis.title.y=element_blank(),
        panel.background=element_rect(fill="white"),
        panel.grid=element_blank())
ggsave("tf_dendro_all.pdf", plot = last_plot(), device = "pdf", path = NULL,
  scale = 1, width = 20, height = 20, units = "cm",
  dpi = 150, limitsize = FALSE)
```

```{r}
#tm.facs.matrix.tfs <- as.matrix(tm.facs.matrix.tfs)
dim(tm.facs.matrix.tfs)
tm.facs.matrix.tfs$cellID <- tm.facs.metadata$cell_ontology_class

tm.facs.tfs.metadata <- dcast(tm.facs.metadata, cell ~ cell_ontology_class)




tm.facs.tfs.ann <- as.data.frame(tm.facs.metadata, row.names = tm.facs.metadata$cell)
rownames(tm.facs.tfs.ann) <- tm.facs.tfs.ann[,1]


tissFACStfs <- CreateSeuratObject(raw.data = tm.facs.matrix.tfs, meta.data = tm.facs.tfs.ann, min.cells = 50)
tissFACStfs <- SetAllIdent(object = tissFACStfs, id = "cell_ontology_class")
prop.table(x = table(tissFACStfs@ident))
cluster.averages <- AverageExpression(object = tissFACStfs, return.seurat = TRUE)



DoHeatmap(object = cluster.averages, genes.use = PCTopGenes(object = tissFACStfs, pc.use = 1, 
    do.balanced = TRUE), group.label.rot = TRUE, group.cex = 0)

DF <- as.matrix(as.data.frame(lapply(cluster.averages, as.numeric)))
heatmap.2(DF)

heatmap.2(cluster.averages)
tail(x = cluster.averages[, 1:5])
TSNEPlot(tissFACStfs, do.label = TRUE, pt.size = 0.5)

tissFACStfs <- ScaleData(tissFACStfs, display.progress = FALSE)
hv.genes <- rownames(tm.facs.matrix.tfs)
tissFACStfs <- RunPCA(tissFACStfs, pcs.print = 0, pc.genes = rownames(tm.facs.matrix.tfs), pcs.compute = 50)
PCElbowPlot(tissFACStfs, num.pc = 50)
tissFACStfs <- FindClusters(tissFACStfs, dims.use = 1:30, print.output = FALSE)
tissFACStfs <- RunTSNE(tissFACStfs, dims.use = 1:30)



tissFACStfs <- NormalizeData(object = tissFACStfs, normalization.method = "LogNormalize", scale.factor = 10000)
hv.genes <- rownames(tm.facs.matrix.tfs)
tissFACStfs <- ScaleData(object = tissFACStfs , genes.use = hv.genes, display.progress = FALSE, 
    vars.to.regress = NULL, do.par = TRUE, num.cores = 1)

tissFACStfs <- RunPCA(object = tissFACStfs, pc.genes = hv.genes, pcs.compute = 100, do.print = TRUE, 
    pcs.print = 1:5, genes.print = 5)


tissFACStfsmeans <- aggregate.Matrix(tissFACStfs@data, groupings = as.factor(tissFACStfs@meta.data$cell_ontology_class), fun = "mean")

tissFACStfs=BuildClusterTree(tissFACStfs,do.reorder = TRUE,reorder.numeric = TRUE)
heatmap.2(as.matrix(as.data.frame(lapply(tissFACStfs@data, as.numeric))))

DF  <-  as.matrix(as.data.frame(lapply(aux, as.numeric)))
DF <- -log(DF,10)
DF[is.na(DF)] <- 0

dd <- dist(DF, method = "euclidean")
hc <- hclust(dd, method = "ward.D2")

heatmap.2(DF)
```



# Subsample data
```{r}
tiss.subsamp <- SubsetData(tiss, max.cells.per.ident = 60, subset.raw = T)
head(unique(tiss.subsamp@meta.data$annotation.2))
```

# Set identity to 

# Function to calculate per-cell type expression averages
```{r}
group_averages <- function(mat, groups){
  group_names = unique(groups)
  means = matrix(0, dim(mat)[1], length(group_names))
  colnames(means) = group_names
  rownames(means) = rownames(mat)
  for(group in group_names){
    means[,group] = Matrix::rowMeans(mat[,groups == group,drop=FALSE])
  }
  means
}

# annotation.2.means <- group_averages(tiss.subsamp@data, tiss@ident)

```

# Correlation + plotting functions
Revision changes: correlations are calculated on the celltype/tissue group averages (arithmetic mean of log1p values),
rather than on single cells from subsampled dataset as in the initial submission.
```{r}
output.tf.cor <- function(data, pval_cutoff, avg_logFC_cutoff, nmarkers.per.celltype, markerfile="TF_markers.csv",height=10){
  
  # read in markers
  markers <- read.csv(markerfile, row.names = 1)
  markers$gene <- rownames(markers)
  markers$avg_logFC <- as.numeric(as.character(markers$avg_logFC))
  markers$p_val <- as.numeric(as.character(markers$p_val))
  
  print(head(markers))
  
  # Get top genes by cell type
  genes.use <- as.character(unique(markers %>% dplyr::group_by(cluster) %>% 
                      filter(gene %in% tf.names, p_val < pval_cutoff, avg_logFC > avg_logFC_cutoff) %>% 
                      dplyr::top_n(n=nmarkers.per.celltype, wt=avg_logFC) %>% 
                      pull(gene)))

  print(paste("genes plotted: ",length(genes.use)))
  # Calculate correlations
  # mat.use <- t(as.matrix(tiss@data[genes.use, ]))
  
  # use cell type averages instead (changed for revision)
  annotation.2.means <- group_averages(data@data, data@ident)
  mat.use <- t(annotation.2.means[genes.use, ])
  tf.cor    <- cor(mat.use)
  
  return(tf.cor)
}


# cluster genes based on expression correlation and plot correlogram
# returns output of barb.cormap, which includes the list of genes in the order they appear in the correlogram
plot.correlogram <- function(data, pval_cutoff, avg_logFC_cutoff, nmarkers.per.celltype, markerfile="TF_markers.csv",height=10){
  
  # read in markers
  markers <- read.csv(markerfile, row.names = 1)
  markers$gene <- rownames(markers)
  markers$avg_logFC <- as.numeric(as.character(markers$avg_logFC))
  markers$p_val <- as.numeric(as.character(markers$p_val))
  
  print(head(markers))
  
  # Get top genes by cell type
  genes.use <- as.character(unique(markers %>% dplyr::group_by(cluster) %>% 
                      filter(gene %in% tf.names, p_val < pval_cutoff, avg_logFC > avg_logFC_cutoff) %>% 
                      dplyr::top_n(n=nmarkers.per.celltype, wt=avg_logFC) %>% 
                      pull(gene)))

  print(paste("genes plotted: ",length(genes.use)))
  # Calculate correlations
  # mat.use <- t(as.matrix(tiss@data[genes.use, ]))
  
  # use cell type averages instead (changed for revision)
  annotation.2.means <- group_averages(data@data, data@ident)
  mat.use <- t(annotation.2.means[genes.use, ])
  tf.cor    <- cor(mat.use)
  print(colnames(tf.cor))
  enrich.score <- dcast(markers, gene ~ cluster, value.var = 'avg_logFC')
  rownames(enrich.score) <- enrich.score[,"gene"]
  enrich.score <- enrich.score[, 2:ncol(enrich.score)]
  # change rownames to the cell type and tissue that each gene is enriched in (ordering by avg_logFC above)
  topIDenriched   <- sapply(colnames(tf.cor), function(x) {
    names(sort(t(enrich.score)[, x],  decreasing = T))[1]}) 
  colnames(tf.cor) <- topIDenriched
  
  correlo.out <- barb.cormap(tf.cor,  'TF_cormap.pdf', height =height, width = height)
  
  # Generate row colors of heatmap corresponding to highest-expressing cell type
  topID.plotorder <- correlo.out[[2]]
  meta.summary <- data@meta.data %>% distinct(tissue, cell_ontology_class, tiss.color, annotation.2)
  print(head(meta.summary))
  ntypes=length(unique(meta.summary$cell_ontology_class))
  tmp <- colorRampPalette(brewer.pal(min(ntypes, 11), 'Paired'))(ntypes)
  annot_colors <- data.frame(annot.colors = tmp, cell_ontology_class = unique(meta.summary$cell_ontology_class))
  write.csv(annot_colors, file=paste0('annotColors.csv'))
  
  print(head(annot_colors))
  meta.summary <- merge(meta.summary, annot_colors, by = 'cell_ontology_class')
  print(head(meta.summary))

  plot.colors <- data.frame(annotation.2 = topID.plotorder)
  print(head(plot.colors))

  plot.colors <- merge(plot.colors, meta.summary %>% select(annotation.2, annot.colors, tiss.color), by = 'annotation.2')
  print((plot.colors))
  plot.colors <- plot.colors[match(topID.plotorder, plot.colors$annotation.2), ]
  print((plot.colors))
  plot.colors$ymin <- 0.1*(0:(nrow(plot.colors)-1))
  plot.colors$ymax <- 0.1*(1:(nrow(plot.colors)))
  print((plot.colors))
  plot.colors$tiss.color <- toupper(plot.colors$tiss.color)
    plot.colors$annot.colors <- toupper(plot.colors$annot.colors)


  require(grDevices)
  pdf( 'TF_cormap_rowcolors_AnnotRight_TissueLeft.pdf', height = 10, width = 4)
    plot(c(0, 2), c(0, max(plot.colors$ymax) + 1), type = "n", xlab = "", ylab = "",
         main = "plot colors")
    rect(0,plot.colors$ymin, 1 , plot.colors$ymax, col = plot.colors$annot.colors, border = NA)
    rect(1,plot.colors$ymin, 2 , plot.colors$ymax, col = plot.colors$tiss.color, border = NA)
  dev.off()

  return(correlo.out)
}

barb.cormap <- function(mat.cor,fname,width=12,height=12,method="complete",cex=0.5,mincor=-1,maxcor=1){
  require(lattice)
  require(cba)
  rowdist <- dist(mat.cor)
  coldist <- dist(mat.cor, by_rows = F)
  hc.cor <- hclust(coldist, method=method)
  hr.cor <- hclust(rowdist, method=method)

  optimal.row <- order.optimal(rowdist,hr.cor$merge)
  optimal.col <- order.optimal(coldist,hc.cor$merge)
 
  ord.row <- optimal.row$order
  ord.col <- optimal.col$order
  
  plt = levelplot(mat.cor[ord.row,ord.col],xlab=NULL,ylab=NULL,
                  at=do.breaks(c(mincor-0.01,maxcor+0.01),19),scales=list(x=list(rot=90),cex=cex),
                  colorkey=list(space="top"),
                  col.regions=colorRampPalette(c("dodgerblue4", "dodgerblue", "white", "lightcoral", "firebrick4"), space="Lab"))
  pdf(fname,width=width,height=height)
    print(plt)
  dev.off()
  
  return(list(rownames(mat.cor[ord.row, ]), colnames(mat.cor[, ord.col]), plt, hc.cor, hr.cor))
}

```

# All cell types TF
Make a list of TFs with cell type specificity to use for correlation analysis.
Version 3 (preparing for final revision, 5/16/18): Seurat FindAllMarkers is faster. FindMarkers now uses much more memory and cannot be easily parallelized w/o crashing (even w/64GB RAM). FindAllMarkers uses 18GB RAM. -->
```{r}
res=FindAllMarkers(tiss.subsamp, genes.use = tf.names,
                   only.pos = T, test.use = "wilcox", min.diff.pct = 0.1, print.bar = F, do.print = T)
write.csv(res, "TF_markers.csv")
```

```{r}
aux <- dcast(res, gene ~ cluster, value.var = "p_val")
rownames(aux) <- aux[,1]
aux <- aux[,-1]
DF  <-  as.matrix(as.data.frame(lapply(aux, as.numeric)))
DF <- -log(DF,10)
DF[is.na(DF)] <- 0

dd <- dist(DF, method = "euclidean")
hc <- hclust(dd, method = "ward.D2")

heatmap.2(DF)
aux <- t(aux)
colnames(aux) <- aux[1,]
aux <- aux[-1,]
```

# compute and plot correlations
```{r}
get_upper_tri <- function(cormat){
    cormat[upper.tri(cormat)]<- NA
    return(cormat)
}

reorder_cormat <- function(cormat){
# Use correlation between variables as distance
dd <- as.dist((1-cormat)/2)
hc <- hclust(dd)
cormat <-cormat[hc$order, hc$order]
}


tf.cor.all <- output.tf.cor(data = tiss.subsamp, pval_cutoff = 10^-3.5, avg_logFC_cutoff = 0.15, nmarkers.per.celltype = 6, height = 30)
reorder_tf.cor.all <- reorder_cormat(tf.cor.all)
upper_tf.cor.all <- get_upper_tri(reorder_tf.cor.all)


melted_tf.cor.all <- melt(upper_tf.cor.all)

# ggplot(data = melted_tf.cor.all, aes(x=Var1, y=Var2, fill=value)) + 
  # geom_tile()

ggplot(data = melted_tf.cor.all, aes(Var2, Var1, fill = value))+
  geom_tile(color = "white")+
  scale_fill_gradient2(low = "white", high = "blue", mid = "white", na.value="white",
   midpoint = 0,  limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal()+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
                                   size = 0, hjust = 1))+
  coord_fixed()+ 
  
  theme(text=element_text(size=4),
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  axis.text.y = element_text(size = rel(1), angle = 0),
  legend.position="none")
ggsave("tf_cor_all.pdf", plot = last_plot(), device = "pdf", path = NULL,
  scale = 1, width = 20, height = 20, units = "cm",
  dpi = 150, limitsize = FALSE)
```


```{r}
correlo.all <- plot.correlogram(data = tiss.subsamp, pval_cutoff = 10^-3.5, avg_logFC_cutoff = 0.15, nmarkers.per.celltype = 6, height = 30)
```


