---
title: "R Notebook"
output: html_notebook
---

```{r}
library(Seurat)
library(cowplot)
library(data.table)
library(dplyr)
library(tidyr)
library(parallel)
library(cba)
library(here)
library(RColorBrewer)
library(tidyverse)
library(openxlsx)
library(gplots)
library(reshape2)

library(ggplot2)
library(ggdendro)
library(scales)
library(wordspace)

library(dendextend)
library(circlize)


library(ontologyIndex)
library(ontologyPlot)
library(ontologySimilarity)
```

# Load Seurat object of data
```{r}
# tm.droplet.matrix = readRDS(here("data-tabula-muris", "TM_droplet_mat.rds"))
# tm.droplet.metadata = read_csv(here("data-tabula-muris", "TM_droplet_metadata.csv"))

tm.facs.matrix = readRDS(here("data-tabula-muris", "TM_facs_mat.rds"))
tm.facs.metadata = read_csv(here("data-tabula-muris", "TM_facs_metadata.csv"))
```

# Load ontology 
```{r}

co = get_ontology('https://raw.githubusercontent.com/obophenotype/cell-ontology/master/cl-basic.obo', extract_tags='everything')

```
Since an ontology is just a collection of `is_a` relationships, the most straightforward distance between nodes will be a function of the two sets of attributes posessed by the nodes.

The number of unshared attributes (the set disjunction) can be easily computed as (union - intersection) of the ancestor sets. In the case of a tree, this is just the tree distance.

```{r}
ont_dist <- function(ontology, id1, id2){
  anc1 = get_ancestors(ontology, id1)
  anc2 = get_ancestors(ontology, id2)
  length(union(anc1, anc2)) - length(intersect(anc1, anc2))
}
```

Since it's not a tree, though, the distance is somewhat larger.

```{r}
plot_common_tree <- function(t1, t2){
  onto_plot(co, terms = intersection_with_descendants(co, "CL:0000548", union(get_ancestors(co, t1), get_ancestors(co, t2))))
}
```

# Load TFs 
```{r}
# (1140 genes from MGI, filtered by GO Term = "DNA binding transcription factor activity", GO ID ?)
tfs      <- read.csv(here('23_tf_analysis','GO_term_summary_20171110_222852.csv'))
tf.names <- as.character( tfs %>% distinct(Symbol) %>% pull(Symbol) )
tf.names <- make.names(tf.names)
length(tf.names)

iegs <- as.character(read.csv(here('00_data_ingest','20_dissociation_genes','genes_affected_by_dissociation_unix.csv'), header = T)[,1])

tf.names <- tf.names[!tf.names %in% iegs]
length(tf.names)
```

# Load Transporter activity 
```{r}
# (1140 genes from MGI, filtered by GO Term = "DNA binding transcription factor activity", GO ID ?)
transporter.activity      <- read.csv(here('23_tf_analysis','GO_term_summary_20180608_203636.csv'))
transporter.activity.names <- as.character( transporter.activity %>% distinct(Symbol) %>% pull(Symbol) )
transporter.activity.names <- make.names(transporter.activity.names)
length(transporter.activity.names)

transporter.activity.names <- transporter.activity.names[!transporter.activity.names %in% iegs]
length(transporter.activity.names)
```

# Load molecular function regulator
```{r}
# (1140 genes from MGI, filtered by GO Term = "DNA binding transcription factor activity", GO ID ?)
mfr      <- read.csv(here('23_tf_analysis','GO_term_summary_20180608_202846.csv'))
mfr.names <- as.character( mfr %>% distinct(Symbol) %>% pull(Symbol) )
mfr.names <- make.names(mfr.names)
length(mfr.names)

mfr.names <- mfr.names[!mfr.names %in% iegs]
length(mfr.names)
```

# Load structural molecular activity
```{r}
# (1140 genes from MGI, filtered by GO Term = "DNA binding transcription factor activity", GO ID ?)
sma      <- read.csv(here('23_tf_analysis','GO_term_summary_20180608_202931.csv'))
sma.names <- as.character( sma %>% distinct(Symbol) %>% pull(Symbol) )
sma.names <- make.names(sma.names)
length(sma.names)

sma.names <- sma.names[!sma.names %in% iegs]
length(sma.names)
```

# Load molecular transducer activity
```{r}
# (1140 genes from MGI, filtered by GO Term = "DNA binding transcription factor activity", GO ID ?)
mta      <- read.csv(here('23_tf_analysis','GO_term_summary_20180608_202922.csv'))
mta.names <- as.character( mta %>% distinct(Symbol) %>% pull(Symbol) )
mta.names <- make.names(mta.names)
length(mta.names)

mta.names <- mta.names[!mta.names %in% iegs]
length(mta.names)
```

# Load catalytic activity
```{r}
# (1140 genes from MGI, filtered by GO Term = "DNA binding transcription factor activity", GO ID ?)
catalytic.activity      <- read.csv(here('23_tf_analysis','GO_term_summary_20180608_202835.csv'))
catalytic.activity.names <- as.character( catalytic.activity %>% distinct(Symbol) %>% pull(Symbol) )
catalytic.activity.names <- make.names(catalytic.activity.names)
length(catalytic.activity.names)

catalytic.activity.names <- catalytic.activity.names[!catalytic.activity.names %in% iegs]
length(catalytic.activity.names)
```

# Load transcription regulator activity
```{r}
# (1140 genes from MGI, filtered by GO Term = "DNA binding transcription factor activity", GO ID ?)
tra      <- read.csv(here('23_tf_analysis','GO_term_summary_20180608_202952.csv'))
tra.names <- as.character( tra %>% distinct(Symbol) %>% pull(Symbol) )
tra.names <- make.names(tra.names)
length(tra.names)

tra.names <- tra.names[!tra.names %in% iegs]
length(tra.names)
```

# Load binding
```{r}
# (1140 genes from MGI, filtered by GO Term = "DNA binding transcription factor activity", GO ID ?)
binding      <- read.csv(here('23_tf_analysis','GO_term_summary_20180608_203841.csv'))
binding.names <- as.character( binding %>% distinct(Symbol) %>% pull(Symbol) )
binding.names <- make.names(binding.names)
length(binding.names)

binding.names <- binding.names[!binding.names %in% iegs]
length(binding.names)
```


# Load cell surface markers 
```{r}
csms      <- read.csv(here('23_tf_analysis','GO_term_summary_20180604_183150.csv'))
csms.names <- as.character( csms %>% distinct(Symbol) %>% pull(Symbol) )
csms.names <- make.names(csms.names)
length(csms.names)

csms.names <- csms.names[!csms.names %in% iegs]
length(csms.names)
```
# Load RNA splicing factors 
```{r}
rna.sfs      <- read.csv(here('23_tf_analysis','GO_term_summary_20171214_190641.csv'))
rna.sfs.names <- as.character( rna.sfs %>% distinct(Symbol) %>% pull(Symbol) )
rna.sfs.names <- make.names(rna.sfs.names)
length(rna.sfs.names)

rna.sfs.names <- rna.sfs.names[!rna.sfs.names %in% iegs]
length(rna.sfs.names)
```


# Calculate global dendogram - FACS with TFs
```{r}

tm.facs.matrix.tfs <- tm.facs.matrix[rownames(tm.facs.matrix) %in% tf.names,]
tm.facs.matrix.tfs <- as.data.frame(t(as.matrix(tm.facs.matrix.tfs)))

tm.facs.matrix.tfs$factors <- as.factor(tm.facs.metadata$cell_ontology_class)

# consider only TFs that are not affected by dissociation
tm.facs.matrix.tfs.cellids <- tm.facs.matrix.tfs %>% 
  group_by(factors) %>%
  summarise_at(vars(colnames(tm.facs.matrix.tfs[1:1016])),funs(sum(., na.rm=TRUE)))


tm.facs.matrix.tfs.cellids.norm = tm.facs.matrix.tfs.cellids
for (i in 1:82){
  jaux = sum(tm.facs.matrix.tfs.cellids[i,2:1017])
  for (j in 2:1017){
    tm.facs.matrix.tfs.cellids.norm[i,j] = tm.facs.matrix.tfs.cellids[i,j]/jaux
  }
}

tm.facs.matrix.tfs.cellids.norm <- as.data.frame(tm.facs.matrix.tfs.cellids.norm)
tm.facs.matrix.tfs.cellids.norm <- subset(tm.facs.matrix.tfs.cellids.norm,!is.na(factors))
rownames(tm.facs.matrix.tfs.cellids.norm) <- tm.facs.matrix.tfs.cellids.norm$factors
tm.facs.matrix.tfs.cellids.norm <- tm.facs.matrix.tfs.cellids.norm[,-1]
dnorm <- dist(tm.facs.matrix.tfs.cellids.norm)
labels(dnorm)

hnorm.tfs <- hclust(dnorm, method = "ward.D2", members = NULL)
```


# Calculate global dendogram - FACS with transporter activity
```{r}

tm.facs.matrix.transporter.activity <- tm.facs.matrix[rownames(tm.facs.matrix) %in% transporter.activity.names,]
tm.facs.matrix.transporter.activity <- as.data.frame(t(as.matrix(tm.facs.matrix.transporter.activity)))

tm.facs.matrix.transporter.activity$factors <- as.factor(tm.facs.metadata$cell_ontology_class)

# consider only TFs that are not affected by dissociation
tm.facs.matrix.transporter.activity.cellids <- tm.facs.matrix.transporter.activity %>% 
  group_by(factors) %>%
  summarise_at(vars(colnames(tm.facs.matrix.transporter.activity[1:1089])),funs(sum(., na.rm=TRUE)))


tm.facs.matrix.transporter.activity.cellids.norm = tm.facs.matrix.transporter.activity.cellids
for (i in 1:82){
  jaux = sum(tm.facs.matrix.transporter.activity.cellids[i,2:1090])
  for (j in 2:1090){
    tm.facs.matrix.transporter.activity.cellids.norm[i,j] = tm.facs.matrix.transporter.activity.cellids[i,j]/jaux
  }
}

tm.facs.matrix.transporter.activity.cellids.norm <- as.data.frame(tm.facs.matrix.transporter.activity.cellids.norm)
tm.facs.matrix.transporter.activity.cellids.norm <- subset(tm.facs.matrix.transporter.activity.cellids.norm,!is.na(factors))
rownames(tm.facs.matrix.transporter.activity.cellids.norm) <- tm.facs.matrix.transporter.activity.cellids.norm$factors
tm.facs.matrix.transporter.activity.cellids.norm <- tm.facs.matrix.transporter.activity.cellids.norm[,-1]
dnorm <- dist(tm.facs.matrix.transporter.activity.cellids.norm)
hnorm.transporter.activity <- hclust(dnorm, method = "ward.D2", members = NULL)
```

# Calculate global dendogram - FACS with molecular function regulator
```{r}

tm.facs.matrix.mfr <- tm.facs.matrix[rownames(tm.facs.matrix) %in% mfr.names,]
tm.facs.matrix.mfr <- as.data.frame(t(as.matrix(tm.facs.matrix.mfr)))

tm.facs.matrix.mfr$factors <- as.factor(tm.facs.metadata$cell_ontology_class)

# consider only TFs that are not affected by dissociation
tm.facs.matrix.mfr.cellids <- tm.facs.matrix.mfr %>% 
  group_by(factors) %>%
  summarise_at(vars(colnames(tm.facs.matrix.mfr[1:1560])),funs(sum(., na.rm=TRUE)))


tm.facs.matrix.mfr.cellids.norm = tm.facs.matrix.mfr.cellids
for (i in 1:82){
  jaux = sum(tm.facs.matrix.mfr.cellids[i,2:1561])
  for (j in 2:1561){
    tm.facs.matrix.mfr.cellids.norm[i,j] = tm.facs.matrix.mfr.cellids[i,j]/jaux
  }
}

tm.facs.matrix.mfr.cellids.norm <- as.data.frame(tm.facs.matrix.mfr.cellids.norm)
tm.facs.matrix.mfr.cellids.norm <- subset(tm.facs.matrix.mfr.cellids.norm,!is.na(factors))
rownames(tm.facs.matrix.mfr.cellids.norm) <- tm.facs.matrix.mfr.cellids.norm$factors
tm.facs.matrix.mfr.cellids.norm <- tm.facs.matrix.mfr.cellids.norm[,-1]
dnorm <- dist(tm.facs.matrix.mfr.cellids.norm)
hnorm.mfr <- hclust(dnorm, method = "ward.D2", members = NULL)
```

# Calculate global dendogram - FACS with structural molecular activity
```{r}

tm.facs.matrix.sma <- tm.facs.matrix[rownames(tm.facs.matrix) %in% sma.names,]
tm.facs.matrix.sma <- as.data.frame(t(as.matrix(tm.facs.matrix.sma)))

tm.facs.matrix.sma$factors <- as.factor(tm.facs.metadata$cell_ontology_class)

# consider only TFs that are not affected by dissociation
tm.facs.matrix.sma.cellids <- tm.facs.matrix.sma %>% 
  group_by(factors) %>%
  summarise_at(vars(colnames(tm.facs.matrix.sma[1:527])),funs(sum(., na.rm=TRUE)))


tm.facs.matrix.sma.cellids.norm = tm.facs.matrix.sma.cellids
for (i in 1:82){
  jaux = sum(tm.facs.matrix.sma.cellids[i,2:528])
  for (j in 2:528){
    tm.facs.matrix.sma.cellids.norm[i,j] = tm.facs.matrix.sma.cellids[i,j]/jaux
  }
}

tm.facs.matrix.sma.cellids.norm <- as.data.frame(tm.facs.matrix.sma.cellids.norm)
tm.facs.matrix.sma.cellids.norm <- subset(tm.facs.matrix.sma.cellids.norm,!is.na(factors))
rownames(tm.facs.matrix.sma.cellids.norm) <- tm.facs.matrix.sma.cellids.norm$factors
tm.facs.matrix.sma.cellids.norm <- tm.facs.matrix.sma.cellids.norm[,-1]
dnorm <- dist(tm.facs.matrix.sma.cellids.norm)
hnorm.sma <- hclust(dnorm, method = "ward.D2", members = NULL)
```

# Calculate global dendogram - FACS with molecular transducer activity
```{r}

tm.facs.matrix.mta <- tm.facs.matrix[rownames(tm.facs.matrix) %in% mta.names,]
tm.facs.matrix.mta <- as.data.frame(t(as.matrix(tm.facs.matrix.mta)))

tm.facs.matrix.mta$factors <- as.factor(tm.facs.metadata$cell_ontology_class)

# consider only TFs that are not affected by dissociation
tm.facs.matrix.mta.cellids <- tm.facs.matrix.mta %>% 
  group_by(factors) %>%
  summarise_at(vars(colnames(tm.facs.matrix.mta[1:2109])),funs(sum(., na.rm=TRUE)))


tm.facs.matrix.mta.cellids.norm = tm.facs.matrix.mta.cellids
for (i in 1:82){
  jaux = sum(tm.facs.matrix.mta.cellids[i,2:2110])
  for (j in 2:2110){
    tm.facs.matrix.mta.cellids.norm[i,j] = tm.facs.matrix.mta.cellids[i,j]/jaux
  }
}

tm.facs.matrix.mta.cellids.norm <- as.data.frame(tm.facs.matrix.mta.cellids.norm)
tm.facs.matrix.mta.cellids.norm <- subset(tm.facs.matrix.mta.cellids.norm,!is.na(factors))
rownames(tm.facs.matrix.mta.cellids.norm) <- tm.facs.matrix.mta.cellids.norm$factors
tm.facs.matrix.mta.cellids.norm <- tm.facs.matrix.mta.cellids.norm[,-1]
dnorm <- dist(tm.facs.matrix.mta.cellids.norm)
hnorm.mta <- hclust(dnorm, method = "ward.D2", members = NULL)
```

# Calculate global dendogram - FACS with catalytic activity
```{r}

tm.facs.matrix.catalytic.activity <- tm.facs.matrix[rownames(tm.facs.matrix) %in% catalytic.activity.names,]
tm.facs.matrix.catalytic.activity <- as.data.frame(t(as.matrix(tm.facs.matrix.catalytic.activity)))

tm.facs.matrix.catalytic.activity$factors <- as.factor(tm.facs.metadata$cell_ontology_class)

# consider only TFs that are not affected by dissociation
tm.facs.matrix.catalytic.activity.cellids <- tm.facs.matrix.catalytic.activity %>% 
  group_by(factors) %>%
  summarise_at(vars(colnames(tm.facs.matrix.catalytic.activity[1:5229])),funs(sum(., na.rm=TRUE)))


tm.facs.matrix.catalytic.activity.cellids.norm = tm.facs.matrix.catalytic.activity.cellids
for (i in 1:82){
  jaux = sum(tm.facs.matrix.catalytic.activity.cellids[i,2:5230])
  for (j in 2:5230){
    tm.facs.matrix.catalytic.activity.cellids.norm[i,j] = tm.facs.matrix.catalytic.activity.cellids[i,j]/jaux
  }
}

tm.facs.matrix.catalytic.activity.cellids.norm <- as.data.frame(tm.facs.matrix.catalytic.activity.cellids.norm)
tm.facs.matrix.catalytic.activity.cellids.norm <- subset(tm.facs.matrix.catalytic.activity.cellids.norm,!is.na(factors))
rownames(tm.facs.matrix.catalytic.activity.cellids.norm) <- tm.facs.matrix.catalytic.activity.cellids.norm$factors
tm.facs.matrix.catalytic.activity.cellids.norm <- tm.facs.matrix.catalytic.activity.cellids.norm[,-1]
dnorm <- dist(tm.facs.matrix.catalytic.activity.cellids.norm)
hnorm.catalytic.activity <- hclust(dnorm, method = "ward.D2", members = NULL)
```

# Calculate global dendogram - FACS with transcription regulator activity
```{r}

tm.facs.matrix.tra <- tm.facs.matrix[rownames(tm.facs.matrix) %in% tra.names,]
tm.facs.matrix.tra <- as.data.frame(t(as.matrix(tm.facs.matrix.tra)))

tm.facs.matrix.tra$factors <- as.factor(tm.facs.metadata$cell_ontology_class)

# consider only TFs that are not affected by dissociation
tm.facs.matrix.tra.cellids <- tm.facs.matrix.tra %>% 
  group_by(factors) %>%
  summarise_at(vars(colnames(tm.facs.matrix.tra[1:1285])),funs(sum(., na.rm=TRUE)))


tm.facs.matrix.tra.cellids.norm = tm.facs.matrix.tra.cellids
for (i in 1:82){
  jaux = sum(tm.facs.matrix.tra.cellids[i,2:1286])
  for (j in 2:1286){
    tm.facs.matrix.tra.cellids.norm[i,j] = tm.facs.matrix.tra.cellids[i,j]/jaux
  }
}

tm.facs.matrix.tra.cellids.norm <- as.data.frame(tm.facs.matrix.tra.cellids.norm)
tm.facs.matrix.tra.cellids.norm <- subset(tm.facs.matrix.tra.cellids.norm,!is.na(factors))
rownames(tm.facs.matrix.tra.cellids.norm) <- tm.facs.matrix.tra.cellids.norm$factors
tm.facs.matrix.tra.cellids.norm <- tm.facs.matrix.tra.cellids.norm[,-1]
dnorm <- dist(tm.facs.matrix.tra.cellids.norm)
hnorm.tra <- hclust(dnorm, method = "ward.D2", members = NULL)
```

# Calculate global dendogram - FACS with binding
```{r}

tm.facs.matrix.binding <- tm.facs.matrix[rownames(tm.facs.matrix) %in% binding.names,]
tm.facs.matrix.binding <- as.data.frame(t(as.matrix(tm.facs.matrix.binding)))

tm.facs.matrix.binding$factors <- as.factor(tm.facs.metadata$cell_ontology_class)

# consider only TFs that are not affected by dissociation
tm.facs.matrix.binding.cellids <- tm.facs.matrix.binding %>% 
  group_by(factors) %>%
  summarise_at(vars(colnames(tm.facs.matrix.binding[1:11546])),funs(sum(., na.rm=TRUE)))


tm.facs.matrix.binding.cellids.norm = tm.facs.matrix.binding.cellids
for (i in 1:82){
  jaux = sum(tm.facs.matrix.binding.cellids[i,2:11547])
  for (j in 2:11547){
    tm.facs.matrix.binding.cellids.norm[i,j] = tm.facs.matrix.binding.cellids[i,j]/jaux
  }
}

tm.facs.matrix.binding.cellids.norm <- as.data.frame(tm.facs.matrix.binding.cellids.norm)
tm.facs.matrix.binding.cellids.norm <- subset(tm.facs.matrix.binding.cellids.norm,!is.na(factors))
rownames(tm.facs.matrix.binding.cellids.norm) <- tm.facs.matrix.binding.cellids.norm$factors
tm.facs.matrix.binding.cellids.norm <- tm.facs.matrix.binding.cellids.norm[,-1]
dnorm <- dist(tm.facs.matrix.binding.cellids.norm)
hnorm.binding <- hclust(dnorm, method = "ward.D2", members = NULL)
```



# Calculate global dendogram - FACS with Cell Surface Markers
```{r}

tm.facs.matrix.csms <- tm.facs.matrix[rownames(tm.facs.matrix) %in% csms.names,]
tm.facs.matrix.csms <- as.data.frame(t(as.matrix(tm.facs.matrix.csms)))

tm.facs.matrix.csms$factors <- as.factor(tm.facs.metadata$cell_ontology_class)

# consider only TFs that are not affected by dissociation
tm.facs.matrix.csms.cellids <- tm.facs.matrix.csms %>%
  group_by(factors) %>%
  summarise_at(vars(colnames(tm.facs.matrix.csms[1:770])),funs(sum(., na.rm=TRUE)))


tm.facs.matrix.csms.cellids.norm = tm.facs.matrix.csms.cellids
for (i in 1:82){
  jaux = sum(tm.facs.matrix.csms.cellids[i,2:771])
  for (j in 2:771){
    tm.facs.matrix.csms.cellids.norm[i,j] = tm.facs.matrix.csms.cellids[i,j]/jaux
  }
}


tm.facs.matrix.csms.cellids.norm <- as.data.frame(tm.facs.matrix.csms.cellids.norm)
tm.facs.matrix.csms.cellids.norm <- subset(tm.facs.matrix.csms.cellids.norm,!is.na(factors))
rownames(tm.facs.matrix.csms.cellids.norm) <- tm.facs.matrix.csms.cellids.norm$factors
tm.facs.matrix.csms.cellids.norm <- tm.facs.matrix.csms.cellids.norm[,-1]
dnorm.csms <- dist(tm.facs.matrix.csms.cellids.norm)
labels(dnorm.csms)

hnorm.csms <- hclust(dnorm.csms, method = "ward.D2", members = NULL)

```


# Calculate global dendogram - FACS with RNAs
```{r}

tm.facs.matrix.rna.sfs <- tm.facs.matrix[rownames(tm.facs.matrix) %in% rna.sfs.names,]
tm.facs.matrix.rna.sfs <- as.data.frame(t(as.matrix(tm.facs.matrix.rna.sfs)))


tm.facs.matrix.rna.sfs$factors <- as.factor(tm.facs.metadata$cell_ontology_class)

# consider only TFs that are not affected by dissociation
tm.facs.matrix.rna.sfs.cellids <- tm.facs.matrix.rna.sfs %>% 
  group_by(factors) %>%
  summarise_at(vars(colnames(tm.facs.matrix.rna.sfs[1:320])),funs(sum(., na.rm=TRUE)))


tm.facs.matrix.rna.sfs.cellids.norm = tm.facs.matrix.rna.sfs.cellids
for (i in 1:130){
  jaux = sum(tm.facs.matrix.rna.sfs.cellids[i,2:321])
  for (j in 2:321){
    tm.facs.matrix.rna.sfs.cellids.norm[i,j] = tm.facs.matrix.rna.sfs.cellids[i,j]/jaux
  }
}

tm.facs.matrix.rna.sfs.cellids.norm <- as.data.frame(tm.facs.matrix.rna.sfs.cellids.norm)
tm.facs.matrix.rna.sfs.cellids.norm <- subset(tm.facs.matrix.rna.sfs.cellids.norm,!is.na(factors))
rownames(tm.facs.matrix.rna.sfs.cellids.norm) <- tm.facs.matrix.rna.sfs.cellids.norm$factors
tm.facs.matrix.rna.sfs.cellids.norm <- tm.facs.matrix.rna.sfs.cellids.norm[,-1]
dnormRNA <- dist(tm.facs.matrix.rna.sfs.cellids.norm)
labels(dnormRNA)

hnorm.rna.sfs <- hclust(dnormRNA, method = "ward.D2", members = NULL)

```

# Calculate ont_dist
```{r}
tm.facs.ids <- unique(tm.facs.metadata$cell_ontology_id[!is.na(tm.facs.metadata$cell_ontology_id)])


sim_mat <- get_sim_grid(ontology=co, term_sets=list(tm.facs.ids))
sim_mat

nids = length(tm.facs.ids)
D = matrix(nrow = nids, ncol = nids)
for (i in 1:nids){
  for(j in 1:nids){
    D[i,j] = ont_dist(co, tm.facs.ids[i], tm.facs.ids[j])
  }
}
rownames(D) = co$name[tm.facs.ids]
colnames(D) = co$name[tm.facs.ids]
d = as.dist(D)

honto <- hclust(d, method = "ward.D2")
```

# Compute and compare dendrogram
# Tanglegram
```{r fig.height = 15, fig.width = 25}

dend1 <- honto %>% as.dendrogram
labels(dend1) <- labels(honto)

dend.tfs <- hnorm.tfs %>% as.dendrogram
dend.ta <- hnorm.transporter.activity %>% as.dendrogram
dend.mfr <- hnorm.mfr %>% as.dendrogram
dend.sma <- hnorm.sma %>% as.dendrogram
dend.mta <- hnorm.mta %>% as.dendrogram
dend.ca <- hnorm.catalytic.activity %>% as.dendrogram
dend.tra <- hnorm.tra %>% as.dendrogram
dend.bind <- hnorm.binding %>% as.dendrogram
dend.rna.sfs <- hnorm.rna.sfs %>% as.dendrogram
dend.csms <- hnorm.csms %>% as.dendrogram

listdend.tfs <- dendlist(dend1, dend.tfs)
listdend.ta <- dendlist(dend1, dend.ta)
listdend.mfr <- dendlist(dend1, dend.mfr)
listdend.sma <- dendlist(dend1, dend.sma)
listdend.mta <- dendlist(dend1, dend.mta)
listdend.ca <- dendlist(dend1, dend.ca)
listdend.tra <- dendlist(dend1, dend.tra)
listdend.bind <- dendlist(dend1, dend.bind)
listdend.rna.sfs <- dendlist(dend1, dend.rna.sfs)
listdend.csms <- dendlist(dend1, dend.csms)


set.seed(3958)

entangle.listdend.tfs <- listdend.tfs %>% untangle(method = "step2side") 
entangle.listdend.ta <- listdend.ta %>% untangle(method = "step2side") 
entangle.listdend.mfr <- listdend.mfr %>% untangle(method = "step2side")  
entangle.listdend.sma <- listdend.sma %>% untangle(method = "step2side") 
entangle.listdend.mta <- listdend.mta %>% untangle(method = "step2side") 
entangle.listdend.ca <- listdend.ca %>% untangle(method = "step2side") 
entangle.listdend.tra <- listdend.tra %>% untangle(method = "step2side")  
entangle.listdend.bind <- listdend.bind %>% untangle(method = "step2side") 
entangle.listdend.rna.sfs <- listdend.rna.sfs %>% untangle(method = "step2side") 
entangle.listdend.csms <- listdend.csms %>% untangle(method = "step2side") 

pdf("entanglements.pdf", width = 50, height = 30)
barplot(c(round(entanglement(entangle.listdend.tfs),2),
          round(entanglement(entangle.listdend.ta),2),
          round(entanglement(entangle.listdend.mfr),2),
          round(entanglement(entangle.listdend.sma),2),
          round(entanglement(entangle.listdend.mta),2),
          round(entanglement(entangle.listdend.ca),2),
          round(entanglement(entangle.listdend.tra),2),
          round(entanglement(entangle.listdend.bind),2),
          round(entanglement(entangle.listdend.rna.sfs),2),
          round(entanglement(entangle.listdend.csms),2)), 
        main="Entanglements", horiz=FALSE,
        xlab = "Entanglement",
        names.arg=c(sprintf("Transcription Factors (n = %s)",length(tf.names)), 
                    sprintf("Transporter Activity (n = %s)",length(transporter.activity.names)),
                    sprintf("Molecular Function Regulator (n = %s)",length(mfr.names)), 
                    sprintf("Structural Molecular Activity (n = %s)",length(sma.names)), 
                    sprintf("Molecular Transducer Activity (n = %s)",length(mta.names)), 
                    sprintf("Catalytic Activity (n = %s)",length(catalytic.activity.names)), 
                    sprintf("Trancription Regulator Activity (n = %s)",length(tra.names)), 
                    sprintf("Binding (n = %s)",length(binding.names)), 
                    sprintf("RNA Splicing Factors (n = %s)",length(rna.sfs.names)), 
                    sprintf("Cell Surface Markers (n = %s)",length(csms.names))),
        col=brewer.pal(10,"Spectral"),
        cex.axis=1.5, cex.names=1.5,
        ylim=c(0,1.1*max(c(round(entanglement(entangle.listdend.tfs),2),
          round(entanglement(entangle.listdend.ta),2),
          round(entanglement(entangle.listdend.mfr),2),
          round(entanglement(entangle.listdend.sma),2),
          round(entanglement(entangle.listdend.mta),2),
          round(entanglement(entangle.listdend.ca),2),
          round(entanglement(entangle.listdend.tra),2),
          round(entanglement(entangle.listdend.bind),2),
          round(entanglement(entangle.listdend.rna.sfs),2),
          round(entanglement(entangle.listdend.csms),2))))
        )
dev.off()

pdf("new_tanglegram_onto_tfs.pdf", width=50, height=30)
x %>% tanglegram(lab.cex = 1.2, lwd = 1,edge.lwd = NULL,common_subtrees_color_branches = TRUE,common_subtrees_color_lines = TRUE,columns_width = c(15, 10, 15),
                      main_left = "Cell Ontology",main_right = "Transcription factors",margin_inner= 22,main = paste("entanglement =", round(entanglement(x), 2)))#,  , dLeaf = -1,type = "r"
dev.off()

pdf("new_tanglegram_onto_csms.pdf", width=50, height=30)
y %>% tanglegram(lab.cex = 1.2, lwd = 1,edge.lwd = NULL,common_subtrees_color_branches = TRUE,common_subtrees_color_lines = TRUE,columns_width = c(15, 10, 15),
                      main_left = "Cell Ontology",main_right = "Cell Surface Markers",margin_inner= 22,main = paste("entanglement =", round(entanglement(y), 2)))#,  , dLeaf = -1,type = "r"
dev.off()

pdf("new_tanglegram_onto_rna_sfs.pdf", width=50, height=30)
z %>% tanglegram(lab.cex = 1.2, lwd = 1,edge.lwd = NULL,common_subtrees_color_branches = TRUE,common_subtrees_color_lines = TRUE,columns_width = c(15, 10, 15),
                      main_left = "Cell Ontology",main_right = "RNA splicing factors",margin_inner= 22,main = paste("entanglement =", round(entanglement(z), 2)))#,  , dLeaf = -1,type = "r"
dev.off()

pdf("new_tanglegram_onto_csms_rna_sfs.pdf", width=50, height=30)
w %>% tanglegram(lab.cex = 1.2, lwd = 1,edge.lwd = NULL,common_subtrees_color_branches = TRUE,common_subtrees_color_lines = TRUE,columns_width = c(15, 10, 15),
                      main_left = "Cell Ontology",main_right = "Cell Surface Markers & RNA splicing factors",margin_inner= 22,main = paste("entanglement =", round(entanglement(w), 2)))#,  , dLeaf = -1,type = "r"
dev.off()

# compare the 2 trees
dend.comp <- all.equal(dend1, dend2, use.edge.length = FALSE, use.tip.label.order = FALSE, use.tip.label = TRUE, use.topology = FALSE, tolerance = .Machine$double.eps^0.5, scale = NULL)

```
Calculate BK distance

```{r}
FM_index(cutree(honto, k=3), cutree(hnorm, k=3)) 
cor_cophenetic(dend1, dend2)
cor_bakers_gamma(dend1, dend2)
cor_bakers_gamma(dend1, dend3)
Bk_plot(dend1, dend1, main = "CORRECT Bk plot \n(based on dendrograms)")
Bk_plot(dend1, dend2, main = "CORRECT Bk plot \n(based on dendrograms)")
Bk_plot(dend1, dend3, main = "CORRECT Bk plot \n(based on dendrograms)")


```



